TLR2 Regulates Hair Follicle Cycle and Regeneration via BMP Signaling

Version 1

Abstract

Abstract

The etiology of hair loss remains enigmatic, and current remedies remain inadequate. Transcriptome analysis of aging hair follicles uncovered changes in immune pathways, including Toll-like receptors (TLRs). Our findings demonstrate that the maintenance of hair follicle homeostasis and the regeneration capacity after damage depends on TLR2 in hair follicle stem cells (HFSCs). In healthy hair follicles, TLR2 is expressed in a cycle-dependent manner and governs HFSCs activation by countering inhibitory BMP signaling. Hair follicles in aging and obesity exhibit a decrease in both TLR2 and its endogenous ligand carboxyethylpyrrole (CEP), a metabolite of polyunsaturated fatty acids. Administration of CEP stimulates hair regeneration through a TLR2-dependent mechanism. These results establish a novel connection between TLR2-mediated innate immunity and HFSC activation, which is pivotal to hair follicle health and the prevention of hair loss and provide new avenues for therapeutic intervention.

Introduction

Hair follicles (HFs), represent one of the best examples of mini-organs with the ability to regenerate throughout life, which, in turn, relies on the proliferation and differentiation of HF stem cells (HFSCs) within hair bulge (Fuchs and Blau, 2020, Sakamoto et al., 2021). The cyclic renewal of HFs is orchestrated by the interplay between inhibitory and stimulatory signals (Plikus et al., 2011). Despite the immune privileged status of HFs, they have a unique microbiome and immune system, including resident macrophages and other immune cells (Fuchs and Blau, 2020, Bertolini et al., 2020, Paus et al., 2003). Components of the HF immune system have been implicated in regulating the HF cycle and its regeneration (Rahmani et al., 2020, Di Domizio et al., 2020). Given their exposure to pathogens, HFs are equipped with innate immune receptors, particularly Toll-like receptors (TLRs), which detect and respond to pathogens by stimulating the secretion of defensins (Di Domizio et al., 2020, Selleri et al., 2007).

Toll-like receptors (TLRs) play a key role in recognizing and responding to either pathogen- or damage-associated molecular patterns, mediating the cytokine response. However, the role of TLRs extends beyond this function, as they have been shown to directly promote tissue regeneration and homeostasis in multiple tissues, particularly in stem and progenitor cells. TLRs regulate hematopoietic and intestinal stem cell renewal, proliferation, and apoptosis (Nagai et al., 2006, Tomchuck et al., 2008). The role of innate immune responses in the tissue-healing benefits of stem cell therapy has been clearly demonstrated (Vagnozzi et al., 2020). Moreover, TLR activation is a critical component of the reprogramming or transdifferentiation of adult cells into pluripotency (Lee et al., 2012), emphasizing the close coordination between innate immunity, cell transformation, and regeneration.

Multiple reports connect altered HFs immunity to hair loss, including a breakdown of immune privilege in alopecia areata (Rahmani et al., 2020). Likewise, androgen, which is tightly linked to TLR activation, was shown to influence the innate immunity of HFs in androgenic alopecia (Sawaya, 2012). The decline of innate immunity processes due to aging or conditions like obesity is widely recognized and these conditions are causatively associated with hair thinning and loss. (Andersen et al., 2016, Shaw et al., 2013, Palmer and Kirkland, 2016, Ghanemi et al., 2020). Alopecia patients often have higher body weight index and weight compared to healthy individuals (Bakry et al., 2014). Increased body weight index is linked to more significant hair loss severity in adults (Goette and Odom, 1976) and a higher prevalence of hair disorders in children and adolescents (Mirmirani and Carpenter, 2014). Mouse models support these findings, showing that activation of innate immunity through pathogen signals might lead to alopecia (Shin et al., 2018) and that high-fat diets inducing obesity cause hair thinning through HFSC depletion (Morinaga et al., 2021).

Our previous studies have shown that activating endothelial TLR2 by endogenous ligands such as CEP (a product of PUFA oxidation) promotes wound healing and tumor angiogenesis (West et al., 2010). Deletion of TLR2 from endothelial cells reduces tumor size by diminishing its vasculature (McCoy et al., 2021). In wounded skin, endothelial TLR2 is crucial for tissue regeneration through increased proangiogenic cytokine secretion (Xiong et al., 2022b). Although PUFAs have been shown to benefit hair growth by extending the anagen phase and promoting cell proliferation and hair shaft elongation (Munkhbayar et al., 2016), the role of innate immunity and in particular, TLR2 in the hair follicle cycle remains unknown.

Using animal models and human cell lines, we show a new function of TLR2 in the hair follicle cycle in homeostasis and hair follicle regeneration in injury. Furthermore, we demonstrate that an endogenously produced PUFA metabolite CEP serves as a TLR2 ligand in the hair bulge, promoting hair regeneration and growth through TLR2. In conditions associated with hair loss, i.e. aging and obesity, both TLR2 and its ligand are substantially depleted in HFs.

Results

TLR2 in HF declines due to aging and obesity

To assess whether and how aging affects hair follicle innate immunity, we analyzed available RNA sequencing data of mouse HFSCs (Doles et al., 2012). Pathway analysis revealed that innate and adaptive immunity, as well as TLR signaling, were among the top dysregulated pathways (Fig. 1A). Notch, JAK-STAT, TGF-β, and Wnt, and other pathways essential for HF regeneration were also altered by aging (Fig. 1A). Notably, the level of Tlr2 mRNA in HFSCs of old mice was two-fold lower compared to young mice (Keyes et al., 2013). In addition, TLR2 at the protein level was substantially lower in 13 months old mice compared to 2 months old mice (Fig. 1B, 1C).

HFSCs downregulate TLR2 in response to stress like a high-fat diet and aging.

A. Dysregulated pathways in old vs young mouse HFSCs. The top pathways are labeled red.

B. Representative confocal images of telogen hair follicles from young and old mice immunostained for TLR2 and CD34 demonstrate decreased TLR2 intensity in HFSC (CD34-positive) of old mice. Scale bars are 50 μm. The middle and right panels show a magnified view of the boxed area. Scale bars are 20 μm.

C. Quantification of TLR2 fluorescent intensity in images from B showing significantly lower TLR2 expression in HFSCs from the old mice. N =6 for each group.

D. GEO2R analysis of published RNA data from sorted follicle populations in the 2nd telogen to anagen transition demonstrates the increased level of TLR2 mRNA accompanied by the activation of TLRs signaling downstream.

E. Representative confocal images showing TLR2 expression in hair follicles from mice fed with a normal diet (ND) or high-fat diet (HFD). CD34 is an HFSC marker. Scale bars are 50 μm. Magnified images demonstrate decreased TLR2 intensity in HFSC (CD34-positive) of mice after HFD. Scale bars are 20 μm.

F. Quantification of TLR2 fluorescent intensity in images from E showing significantly lower TLR2 expression in HFSCs from HFD-fed mice. N = 7 and 6 for ND and HFD groups respectively. AU, arbitrary unit.

G. TLR2 mRNA expression in HFSCs from mice fed with ND or HFD for 4 days or 3 months. Data regenerated from published RNA sequencing dataset GSE131958. N = 3 for each group.

All bar graphs are mean ± s.e.m. Non-parametric Mann-Whitney test (C, G) or unpaired two-tailed t-test (F) was used to determine statistical difference.

At the same time, the normal hair cycle is marked by an increase in TLR2 mRNA levels prior to HFSC activation during telogen, according to the analysis of existing RNAseq data (Fig. 1D) (Greco et al., 2009). TLR2 mRNA levels are highest among other TLRs, with TLR1 and TLR4 mRNA showing the opposite pattern. TLR6 may act as a co-receptor for TLR2 as its expression pattern is similar. While the downstream TLR2 signaling molecules, IRAK1 and MYD88, are also upregulated, mirroring the TLR2, IRAK1 inhibitor TOLLIP is suppressed. Together, the results suggest the critical regulatory role of the entire TLR2/TLR6 pathway in the hair follicle cycle.

High-fat diet-induced obesity causes hair thinning and subsequent loss (Morinaga et al., 2021). In our model, a high-fat diet causes a nearly 2-fold decline in TLR2 protein level in HFSCs, compared to normal diet-fed mice (Fig. 1E, 1F). Further, RNA sequencing data reveal that 3 months of a high-fat diet is sufficient to reduce TLR2 levels in HFSCs by more than 2-fold compared to control mice (Fig. 1G) (Morinaga et al., 2021). Thus, our results and analyses of existing datasets demonstrate that conditions causatively associated with hair thinning and loss, such as aging and obesity, result in a dramatic depletion of TLR2 in HFSCs suggesting a possible regulatory role for TLR2 in hair follicles.

TLR2 is upregulated during HFSC activation

The expression of TLR2 during a normal hair cycle was assessed using a previously characterized TLR2-GFP reporter mouse (Fig. 2), one of the best tools for the analysis of TLRs in vivo (Price et al., 2018). The correlation between the reporter and TLR2 protein expression was confirmed (Supplementary fig. 1A, 1B). The hair follicle cycle was verified by H&E staining (Supplementary fig. 1C). During telogen, a dormant stage for HFSCs, TLR2 was found in the bulge, secondary hair germ (sHG), dermal papilla (DP), and outer root sheath (ORS) (Fig. 2A). During regenerative anagen, TLR2 expression was detected in the DP and all sHG-derived progenitor cells (Fig. 2B, 2C) and in quiescent bulge stem cells (Fig. 2D, 2E). All cells derived from bulge stem cells (ORS) and sHG (hair shaft and inner root sheath, IRS) were positive for TLR2 (Fig. 2C). In catagen, TLR2 was abundant in the new bulge and sHG formed from ORS cells (Hsu et al., 2011) (Fig. 2F, 2G). While TLR2 was expressed in early sHG lineage (including IRS) (Fig. 2C), it was absent in mature (Fig. 2D, 2E) and regressing (Fig. 2F) IRS. This shows that TLR2 is abundant in stem cells but declines upon differentiation. The second telogen’s old and new bulges expressed TLR2 (Fig. 2H). TLR2 was present in sHG and DP during the second telogen (Fig. 2H) and increased in the late (competent) telogen compared to the early (refractory) telogen (Supplementary fig. 1D, 1E). The highest level of TLR2 occurred during active anagen compared to quiescent telogen and catagen (Fig. 2I). Quantitative PCR revealed that the Tlr2 mRNA level in HFSCs was 5- and 2.3-fold higher in anagen than in telogen and catagen, respectively (Fig.2J). Notably, analysis of existing RNA sequencing data using FACS-sorted cells (Lorz et al., 2010) confirmed that TLR2 expression was significantly higher in HFSCs than in epidermal or non-stem cells (Fig. 2K). Thus, TLR2 is enriched in HFSCs, and its expression increases during activation.

TLR2 is enriched in HFSCs and is upregulated during HFSC activation.

TLR2-GFP reporter mouse skin sections were immunostained with anti-GFP to assess TLR2 expression in the hair follicles.

A. Representative confocal images of P21 first telogen hair follicle immunostained for TLR2-GFP, CD34 (bulge stem cells), P-cad (secondary hair germ (sHG)), and DAPI (nuclei). The green color in the surface rendering panel represents TLR2 expression, and other surfaces show co-localization between TLR2 and specific markers. TLR2 is present in bulge, sHG, and dermal papilla (DP) cells. P represents postnatal days. Scale bar is 10 μm.

B. TLR2-GFP in P28 anagen was co-immunostained with CD49f of basement membrane outlining the DP Scale bar is 10 μm.

C. TLR2 is co-localized to the sHG lineage (P-cad+ layers), DP, and outer root sheath (ORS) lineage. Scale bar is 20 μm.

D. TLR2-GFP in P28 anagen was co-immunostained with CD34 in old bulge (D) and Ker5 in ORS (E) revealing TLR2 localization to the old bulge, ORS, but not inner root sheath (IRS). Scale bars are 20 μm.

F. Co-immunostaining of TLR2-GFP in P38 catagen hair follicle with Ker5 in ORS lineage cells showing co-localization of TLR2 with ORS and bulge. Scale bar is 20 μm.

G. P41 late catagen hair follicle immunostained for TLR2 and CD34 showing co-localization of TLR2 to the old bulge, new bulge, sHG, and DP Scale bar is 20 μm.

H. P53 second telogen hair follicle immunostained for TLR2, CD34, and P-cad reveals co-localization of TLR2 to the bulge, sHG, and DP Scale bar is 20 μm.

I. Quantification of TLR2 fluorescent intensity in bulge cells at different phases showing TLR2 upregulation in anagen. N = 3 for each group.

J. QPCR analysis of Tlr2 mRNA expression in FACS-purified mouse HFSCs in anagen, telogen, and catagen. N = 3 or 4 per group.

K. QPCR analysis of Tlr2 mRNA expression in mouse epidermal cells and FACS-purified HFSCs showed significantly higher Tlr2 expression in HFSCs compared with raw epidermal cells. N = 6 mice per group.

All bar graphs are mean ± s.e.m. Two-tailed unpaired t-test (K) or Kruskal-Wallis test with Dunn’s post hoc test (I, J) was used to determine statistical difference.

Deletion of TLR2 in HFSCs delays anagen onset in the normal hair cycle

To address the role of TLR2 in the hair cycle, we generated an HFSC-specific inducible Tlr2 knockout mouse line (TLR2HFSC-KO) and deleted Tlr2 during the first postnatal telogen (Xiong et al., 2022b). In TLR2HFSC-KO mice, the telogen phase was substantially prolonged compared to control (TLR2lox/lox) mice, as summarized in the schematic (Fig. 3A). Melanogenesis and anagen onset are tightly coupled (Muller-Rover et al., 2001). Thus, the pink skin color at P21 marks the first postnatal telogen. The onset of anagen in control mice was indicated by the change in skin color to gray or black at P26 in control mice. TLR2HFSC-KO mice at this age did not enter anagen, as evidenced by the delayed darkening of their skin (Fig. 3A-3C). This was further confirmed by skin section analysis at P21, P26, and P35 (Fig. 3D). At P26, control mice displayed pigmented anagen hair follicles with enlarged bulbs located deeply in the hypodermis, while nearly all follicles of TLR2HFSC-KO mice were ∼5 fold shorter and remained in the dermis on top of adipose tissue, a characteristic of telogen. On day p35, we observe partial entrance into anagen in TLR2HFSC-KO skin while the skin color of TLR2HFSC-KO mice remains pink (Fig. 3D-3F). HFSCs were activated as early as at P24 in control mice based on positive Ki67 staining in sHG and bulge region (Fig. 3G), while most cells in TLR2HFSC-KO sHG (Fig. 3H) and bulge (Fig. 3I) remained quiescent. At P25, control mice exhibited a large cluster of P-cad+ cells encapsulating DP within the transformed sHG (Fig.3K), whereas the sHG of TLR2HFSC-KO mice remained small and inactive (Fig. 3J, 3K and Supplementary fig. 2). Despite the substantial delay in anagen onset, the morphology of hair follicles and expression of established HFSC markers, including Ker15, CD34, and Sox9, were normal in TLR2HFSC-KO mice (Supplementary fig. 3). Thus, TLR2 in HFSCs is essential for HFSC activation and progression of the hair cycle.

Deletion of TLR2 in hair follicle stem cells delays anagen onset.

A. Schematic of RU486-mediated Cre induction and dorsal skin pigmentation change (gradient bars) in TLR2lox/lox and TLR2HFSC-KO mice.

B. Representative images of shaved TLR2lox/lox and TLR2HFSC-KO mice showing different phases of the hair cycle. The TLR2lox/lox mouse transitions from telogen (pink skin) to anagen (grey/black skin) at P26 and a full hair coat is developed by P35. The TLR2HFSC-KO mouse exhibits a prolonged telogen (P21-P30-P35). Representative images of at least 10 mice in each group.

C. Bar graph showing the length of first postnatal telogen starting from P21 measured by skin color change from B. N = 10 per group.

D. Representative H&E staining of dorsal skin at indicated time points showing prolonged telogen in TLR2HFSC-KO mice. Scale bars are 50 μm.

E. The length of hair follicles at P26 from images in D. 50 hair follicles from 3 mice per group were used for quantification.

F. Percentages of telogen or anagen hair follicles at P26 from D. N = 4 mice per group.

G. Representative confocal images of P21 and P24 first telogen hair follicles from TLR2lox/lox and TLR2HFSC-KO mice immunostained for CD34, Ki67, and DAPI. Stars label the hair shaft. Scale bars are 20 μm.

H, I. Quantification of images in G shows a diminished number of Ki67+ cells in sHG (H) and in CD34+ bulge (I) in TLR2HFSC-KO mice compared to TLR2lox/lox at P24. N = 6 and 4 mice for TLR2lox/lox and TLR2HFSC-KO group respectively.

J. Representative confocal images of P21 and P25 dorsal skin sections from TLR2lox/lox and TLR2HFSC-KO mice immunostained for P-cad and DAPI showing changes in the size of sHG. Scale bars are 20 μm.

K. Quantification of sHG size in panel K shows enlarged sHG in TLR2lox/lox mice compared with TLR2HFSC-KO mice. N=4 mice for p 25 TLR2lox/lox, and N=5 mice for TLR2HFSC-KO. Statistical significance was determined using a non-parametric Mann-Whitney test. All data are mean ± s.e.m.

TLR2 regulates HFSC activation by interacting with BMP signaling pathway

The relationship between WNT and BMP signaling is central to the cyclic growth of hair follicles (Plikus et al., 2008). Anagen initiation is triggered by Wnt/β-catenin activation, while BMP signaling suppresses HFSC activation and its reduction is necessary for HFSC activation. Indeed, during the early (refractory) phase of the second telogen, HFSCs exhibit elevated BMP signaling as evidenced by high levels of BMP7 protein and pSMAD1/5/9, downstream targets of BMP signaling, compared to the late (competent) phase (Fig. 4A-4D). In contrast, BMP7 and its effectors, ID1 and ID2, are decreased during the late telogen based on our analysis of existing RNAseq datasets (Greco et al., 2009) (Supplementary Fig. 4A).

TLR2 interacts with BMP pathway to regulate the hair cycle.

A. Representative confocal images of BMP7 staining in hair follicles of dorsal skin in early (p46) and late (p69) second telogen. Scale bars are 10 μm.

B. Quantification of BMP7 fluorescent intensity from A showing diminished BMP7 expression during the second telogen from the early to late phases. N = 4 per group.

C. Representative confocal images of pSMAD1/5/9 staining in hair follicles of dorsal skin in early (p46) and late (p69) second telogen. Scale bars are 10 μm.

D. Quantification of pSMAD1/5/9+ positive cells in CD34+ bulge stem cells demonstrates a decrease of pSmad1/5/9 expression in late telogen. N = 4 per group.

E. QPCR analysis reveals dysregulation of BMP singling molecules in HFSCs lacking TLR2. N=4 mice for Control and BMP2, N=3 mice for BMP7 and BMPr1A.

F. Representative confocal images of BMP7 staining in hair follicles from TLR2lox/lox or TLR2HFSC-KO mice. Scale bars are 10 μm. Stars label hair shaft.

G. Quantification of BMP7 fluorescent intensity from F showing higher BMP7 expression in TLR2HFSC-KO mice. N = 4 per group.

H. P21 and P24 dorsal skin sections from TLR2lox/lox and TLR2HFSC-KO mice immunostained for CD34, pSmad1/5/9, and DAPI. Scale bars are 10 μm.

I. Quantification of pSmad1/5/9+ cells in CD34+ bulge stem cells in P24 dorsal skin from H. N = 4 and 5 for TLR2lox/lox and TLR2HFSC-KO respectively.

J. Representative confocal images of dorsal skin sections from TLR2HFSC-KO mice treated with BSA or noggin immunostained for CD34, pSmad1/5/9, and DAPI. Star labels the hair shaft. Scale bars are 10 μm.

K. Quantification of pSmad1/5/9+ cells in CD34+ bulge stem cells from images in J. N = 5 per group.

L. Immunostaining for Ki67 and DAPI in dorsal skin sections from TLR2HFSC-KO mice treated with BSA or noggin. Scale bars are 10 μm.

M. Quantification of images in L showing an increase in Ki67+ cells in sHG of noggin-treated compared to BSA-treated TLR2HFSC-KO dorsal skin. N = 5 per group.

N. Representative confocal images of Ki67 and DAPI immunostaining of dorsal skin sections from TLR2HFSC-KO mice treated with BSA or noggin. Arrows point to hair follicles with Ki67+ cells in the sHG. Scale bars are 20 μm.

O. Quantification of images in N showing percentages of hair follicles with Ki67+ cells in sHG. N = 5 per group.

P. BSA- or noggin-treated TLR2HFSC-KO mouse dorsal skin immunostained for P-cad and DAPI. The dashed line outlines the sHG. Scale bars are 10 μm.

Q. Bar graph showing significantly larger sHG in noggin-treated TLR2HFSC-KO mice. N = 5 per group.

Mann-Whitney test was used to determine the statistical significance. All data are mean ± s.e.m.

To assess the possible connection between the TLR2 signaling and BMP pathway in human cells, we activated TLR2 and BMP signaling in human epidermal keratinocytes (NHEK) using a canonical TLR2 agonist (Pam3CSK4) and BMP4, respectively. As anticipated, BMP4 promoted the phosphorylation of its downstream target SMAD1/5/9. However, simultaneous co-activation of TLR2 diminished BMP4 signaling (Supplementary fig. 4B, 4C).

Likewise, stimulation of human HFSC with canonical TLR2 agonist Pam3CSK4 promoted cell proliferation by 1.5-fold compared to controls. Notably, this effect was diminished in the presence of a TLR2-blocking antibody (Supplementary fig. 4D, 4E). These results reveal that TLR2 activation on human HFSC augments their proliferation.

To gain a deeper understanding of TLR2’s role in HFSC activation, we profiled the transcriptome of HFSCs lacking TLR2 expression (Supplementary fig. 4F). The results showed that TLR2 deletion dysregulated 486 genes, many of which were involved in both the hair cycle and innate immunity. The most affected pathways included innate immunity response and TLR2 signaling, with its downstream target NF-kappaB (Supplementary fig. 4G). This profile somewhat resembles changes observed in aging models (Fig. 1A).

BMP pathway is altered in TLR2 KO HFSCs

Since TLR2 suppresses BMP signaling and promotes HFSC proliferation, we assessed whether the delayed anagen in TLR2HFSC-KO mice might be associated with the BMP pathway. QPCR analysis reveals that several key components of the BMP pathway were dysregulated in HFSCs lacking TLR2 (Fig.4E). Among those, the most notable changes were observed for BMP7, which was upregulated by ∼4-fold in TLR2-null HFSCs compared to controls (Fig. 4E). This was substantiated by co-staining of tissue sections for BMP7 and CD34, which demonstrated a ∼2-fold increase in BMP7 on HFSCs of TLR2HFSC-KO mice as compared to the control (Fig. 4F, 4G). Activation of BMP signaling was assessed by pSMAD1/5/9 positive staining in HF. Quantification revealed a ∼15-fold higher level of pSMAD1/5/9 in TLR2-null mice (TLR2KO) as compared to controls (wild type) (Supplementary fig. H,4I). The significant increase in BMP signaling observed was attributed to the absence of TLR2 in hair follicle stem cells (HFSCs). This was evidenced by a comparable 14-fold increase in BMP signaling in follicles of TLR2HFSC-KO mice compared to control during the first postnatal telogen phase, thereby ensuring the preservation of follicles in the dormant telogen stage (as shown in Fig. 4H and 4I). Simultaneously, the Wnt signaling and β-catenin stabilization within HFSCs, known to trigger their activation (Deschene et al., 2014), remained unchanged between control and TLR2HFSC-KO mice (as shown in Supplementary Fig. 4J).

BMP antagonist rescues defects caused by the lack of HFSCs TLR2

To demonstrate that an altered BMP pathway is, indeed, responsible for the phenotype of TLR2 knockout in HFSCs, we utilized intradermal injection of noggin, a well-known inhibitor of BMP signaling (Botchkarev et al., 2001, Botchkarev et al., 1999), to block the upregulated BMP signaling in TLR2HFSC-KO mice. As a result, noggin injection diminished activation of BMP signaling by >10-fold in TLR2HFSC-KO mice as assessed by pSMAD1/5/9 staining of HF (Fig. 4J, 4K). Moreover, noggin promoted activation of TLR2HFSC-KO HFs while the HFs in BSA-treated TLR2HFSC-KO mice remained quiescent. Noggin treatment of TLR2HFSC-KO mice dramatically upregulated cell proliferation within sHG as evidenced by Ki67+ cells (Fig. 4L, 4M), promoting ∼2.5 fold increase in activated follicles (Fig. 4N, 4O), thereby contributing to nearly 2-fold larger sHG (Fig. 4P, 4Q) as compared to BSA-treated TLR2HFSC-KO mice. Thus, curbing suppressive BMP signaling in TLR2HFSC-KO mice can reactivate their HFs, demonstrating a causative connection between TLR2 and BMP pathways in the hair cycle.

HFSC TLR2 governs hair regeneration upon injury

High expression of TLR2 and its critical role in HFSC activation during the hair cycle prompted us to test the role of HFSCs TLR2 in an injury model where cells are more likely to be exposed to TLR2 ligands. First, we compared TLR2 levels in HFs in wounded and healthy skin using TLR2-GFP reporter mouse (Fig. 5A). In healthy skin, HFSCs upregulated TLR2 during their transition from middle to late telogen (day 5 to day 10) (Fig. 5B upper panels, and grey bars in fig. 5C), consistent with RNA sequencing results (Greco et al., 2009). This increase in TLR2 precedes HFSCs activation during the normal cycle. However, in wound HFSCs, TLR2 was upregulated immediately after an injury resulting in 1.5-fold higher expression compared to normal unwounded skin (Fig. 5B, 5C).

HFSC TLR2 is crucial for wound-induced hair follicle regeneration.

A. Schematic of wound healing assay using TLR2-GFP reporter mouse. Full-thickness wounds on the dorsal skin of TLR2-GFP mice were created. Normal unwounded skin and the skin adjacent to the wound were harvested at different time points.

B. Representative confocal images of normal and wounded skin from TLR2-GFP mice at different time points post-injury immunostained for TLR2-GFP and CD34. Scale bars are 20 μm.

C. Quantification of TLR2 fluorescent intensity per bulge cell in hair follicles from B shows increased TLR2 level in hair follicles from wounded skin as compared to normal skin. N=3 for Day 1 and Day 5, and N=4 for Day 10 per group.

D. Representative photographs showing hair regeneration on the dorsal skin and inner skin flaps at indicated time post-injury in WT and TLR2 global knockout (TLR2KO) mice. Diminished hair growth around the wound is apparent in TLR2KO skin from day 14 through 21 post-injury compared to WT skin. The inner skin flaps from TLR2KO at day 14 post-injury show an absence of pigmented hair bulbs and skin pigmentation. Scale bars are 5mm for dorsal skin and 1mm for inner skin flaps.

E. Quantification of pigmented dorsal skin area around the wound from images in D shows diminished pigmentation in TLR2KO skin compared with WT skin at all time points post-injury. N=4 per group.

F. Representative confocal images of skin adjacent to wound immunostained for Ki67 and DAPI. Scale bars are 50 μm.

G. Bar graph showing diminished Ki67 fluorescent intensity in the skin adjacent to wound in TLR2KO mouse compared to WT mouse from images in F. N = 4 and 7 for WT and TLR2KO respectively.

H. Representative confocal images of skin adjacent to wound immunostained for Ker17 and DAPI. Scale bars are 100 μm.

I. Quantification of hair follicle numbers from images in H reveals a significant decrease in regenerated hair follicles in TLR2KO skin compared with WT skin. N = 5 and 7 for WT and TLR2KO respectively.

J. Representative photographs showing a lack of hair regeneration and skin pigmentation around the wound on the dorsal skin of TLR2HFSC-KO mice compared with TLR2lox/lox mice on day 17, day 21, and day 24 post-injury. Scale bars are 2mm.

K. Quantification of pigmented skin area around the wound during 14-28 days post-injury showing significantly smaller pigmented skin area in TLR2HFSC-KO mice compared with TLR2lox/lox mice. N = 4 per group.

L. Representative confocal images of wounded skin from TLR2lox/lox and TLR2HFSC- KO mice stained for CD34 and pSmad1/5/9. Scale bars are 10 μm.

M. Quantification of images from L showing more pSmad1/5/9+ cells in TLR2HFSC-KO wounded skin. N = 3 per group.

Mann-Whitney test was used to determine the statistical significance. All data are mean ± s.e.m.

Hair regeneration after injury represents a substantial part of the healing process (Wang et al., 2017, Chen et al., 2015, Abbasi and Biernaskie, 2019). We next assessed the role of TLR2 using age- and gender-matched WT and TLR2KO mice. The lack of TLR2 visibly impaired hair regeneration after wound healing (Fig.5D). On day 14 post-injury, HFs in WT mice entered precocious anagen judged by a spot of pigmented skin, which, on day 21 developed into a black hair patch (Fig. 5D). In contrast, the follicles of TLR2KO mice remained quiescent lacking regenerated HFs around wounds even after 21 days post-injury (Fig. 5D, 5E). At this point, pigmented skin area in WT was ∼9-fold larger than in TLR2KO mice. Skin flaps showed substantial pigmentation and growing hair bulbs around wounds in WT, indicative of active anagen. In contrast, the TLR2KO skin flap was devoid of pigmentation, consistent with telogen (inner skin flap in fig. 5D). Ki67 staining confirmed an increase in HFs activation in WT but not in TLR2KO skin (Fig. 5F, 5G). The resulting density of regenerated HFs based on Ker17 staining in WT was 2-fold higher than in TLR2KO mice (Fig. 5H, 5I). Most importantly, this effect was dependent on TLR2, specifically on HFSCs, since TLR2HFSC-KO mice exhibited a similar phenotype with a dramatic reduction in pigmentation and hair growth compared to control mice (Fig. 5J, 5K). The upregulation of pSmad1/5/9 in TLR2HFSC-KO wounds compared to controls demonstrates that similar to the HF cycle scenario, increased BMP signaling might contribute to diminished HF regeneration (Fig. 5L, 5M). Thus, the TLR2-BMP axis in HFSCs governs HF regeneration after injury.

Endogenous ligand promotes hair regeneration via TLR2 on HFSCs

One of the most important endogenous ligands for TLR2 is CEP, which is a naturally occurring product of PUFA oxidation shown to be accumulated during inflammation and wound healing (West et al., 2010, Xiong et al., 2022a). Healthy tissues are typically devoid of this product, which is mainly associated with inflammation and pathologies (Yakubenko et al., 2018). However, in contrast to other tissues, healthy HFs exhibited high levels of CEP accumulation (Fig. 6A). During anagen, CEP is present within the proximal part of the follicle, while in telogen the entire follicle is encased by this PUFA metabolite (Fig. 6B-6E). Generation of CEP from PUFA is directly aided by myeloperoxidase (MPO) (Yakubenko et al., 2018, Xiong et al., 2022a). MPO is present in abundance in sebaceous glands, possibly as a part of immune defense (Supplementary fig. 5A). Even more surprising, in contrast to other organs and tissues, CEP in HFs is substantially depleted with age (Fig. 6F, 6G), and this decline coincides with the reduction in the regenerative potential of HFs. This is likely due to decreased level of MPO during aging (Supplementary fig. 5B, 5C).

Oxidation-dependent TLR2 ligand CEP is present in hair follicles and promotes hair regeneration via HFSC TLR2.

A. Representative images of H&E and CEP immunostaining of consecutive skin sections from WT anagen mouse. Scale bars are 1 mm.

B. Representative confocal images of P5 WT whole-mount skin immunostained for CEP and Ker17. The merged image shows the co-localization of CEP to anagen hair follicles (Ker17+). Scale bar is 200 μm.

C. Longitudinal and cross-sections of anagen and telogen hair follicles from WT mice immunostained for CEP and Ker17. The lower left panel shows a magnified view of the boxed area. Scale bars are 100 μm for anagen, 50 μm for telogen.

D. Quantification of CEP fluorescent intensity at a different distance from the root of anagen hair follicles in longitudinal and cross-sections immunostaining images in images from C. A gradual decrease in CEP levels is observed from the proximal to the distal part of anagen hair follicles. N= 50 follicles from 3 per group.

E. Line chart showing a sharp decrease of CEP fluorescent intensity with the distance from HF in telogen (from the lower right panel in C). N= 10 follicles from 3 per group.

F. Representative confocal images of telogen hair follicles from young and old mice immunostained for CEP and Ker17. Scale bars are 20 μm.

G. Quantification of CEP fluorescent intensity from images in F. N = 6 mice per group.

H. Representative photographs of dorsal skin (two left panels) and inner skin flaps (two right panels) from WT and TLR2KO mice after irradiation and bone-marrow transplantation of WT bone marrow demonstrate an increased number of pigmented hair bulbs and skin pigmentation around wounds in CEP-treated wounds compared to control in WT mice with no differences in TLR2KO transplanted with WT bone marrow. Scale bars are 1 mm for the dorsal skin and 500 μm for the inner skin flap.

I. Quantitative results from H show an increased density of hair follicles upon CEP application around wounds of WT> WT transplanted mice with no changes in WT> TLR2KO mice. N = 5 for each group.

J. Representative photographs of dorsal skin (upper panels) and inner skin flaps (lower panels) from TLR2lox/lox and TLR2HFSC-KO mice treated with CEP show a lack of pigmentation around TLR2HFSC-KO wounds compared with TLR2lox/lox wounds treated with CEP. The inner skin flap of TLR2HFSC-KO mice demonstrates an absence of pigmented hair bulbs after the CEP treatment Scale bars are 3 mm.

K. Representative confocal images of skin adjacent to wound immunostained for Ker17. Scale bars are 100 μm.

L. Quantification of hair follicle numbers in images from K reveals a significant decrease in regenerated hair follicles in TLR2HFSC-KO skin compared with TLR2lox/lox skin.

N=7 for TLR2lox/lox N = 4 for TLR2HFSC-KO.

M. Representative confocal images of skin adjacent to wound immunostained for Ki67. Scale bars are 50 μm.

N. Bar graph showing Ki67 fluorescent intensity in the skin adjacent to wound from images in M. N=7 for TLR2lox/lox N = 4 for TLR2HFSC-KO.

O. Representative microphotographs of primary keratinocytes isolated from WT or TLR2KO mouse skin co-cultured with CEP or control (PBS or BSA). Representative images from at least three independent assays are shown. Scale bar 50 µm.

P. Cell proliferation of primary keratinocytes in O indicates increased proliferation by CEP in WT but not in TLR2KO keratinocytes. N=3 independent experiments.

Q. QPCR analyses of Nfkb2, Il1b, and Il6 mRNA levels in FACS-purified mouse HFSCs treated with BSA control or CEP. N = 3 per group.

R. QPCR analyses of Bmp7 mRNA levels in FACS-purified mouse HFSCs treated with BSA control or CEP. N = 3 per group.

S. Summary of the main findings of this study.

Unpaired t-test (G, P) or Mann-Whitney test (I, L, N, Q, R) was used to determine the statistical significance. All data are mean ± s.e.m.

The connection between CEP levels and hair thinning and loss in aging prompted us to test whether exogenous CEP can activate TLR2 in HFSCs and stimulate their proliferation. Our in-vitro experiments revealed that CEP increases the proliferation of human HFSC in a TLR2-dependent manner since the blockade of TLR2 abrogates the CEP effect (Supplementary fig. 5D, 5E). In another model, CEP promotes cell proliferation of human hair follicle dermal papilla cells (HFDPC) by ∼2-fold compared to control (Supplementary fig. 5F).

Next, we show that CEP promotes hair regeneration in injury in a TLR2-dependent manner. CEP administration promoted HF regeneration in WT wounds. However, it was ineffective in global TLR2KO mice (Supplementary fig. 5H). CEP promoted a 55% increase in the number of HF and cell proliferation in WT wounds, and at the same time, there was no effect in TLR2KO wounds (Supplementary fig. 5I-5J).

To ensure independence from immune cells, WT and TLR2KO mice were irradiated and transplanted with WT bone marrow prior to wounding. Applying CEP on wounds in WT/WT chimeras promoted cell proliferation, thereby dramatically increasing the density of hair follicles (Fig.6H, 6I, Supplementary fig. 5G). At the same time, CEP was not effective in TLR2KO/WT mice (Fig.6H, 6I, Supplementary fig. 5G), demonstrating the TLR2-dependent mechanism.

These CEP effects were mediated by TLR2 on HFSCs. In control mice, CEP effectively initiated regeneration of HFs around the wound (Fig.6J), resulting in ∼3-fold higher density of HF (Fig.6K, 6L) and dramatic acceleration of cell proliferation by >10 fold (Fig.6M, 6N) as compared to TLR2HFSC-KO mice where CEP was mainly ineffective. A similar stimulatory effect of CEP was observed in a primary keratinocyte culture (Oshimori and Fuchs, 2012). CEP dramatically promoted WT but not TLR2KO keratinocyte proliferation (Fig. 6O, 6P). CEP was an effective stimulator of TLR2 signaling as judged by augmented Nfkb2, Il1β, and Il6 expression in HFSCs upon treatment with CEP (Fig. 6Q). Consistent with the key role of BMP signaling in TLR2-dependent HF regeneration, CEP treatment suppressed inhibitory Bmp7 expression by ∼2.5 fold (Fig. 6R), demonstrating that endogenous and natural TLR2 ligand can counteract an inhibitory effect of BMP7 to stimulate HFSCs activation (Fig. 6S).

Discussion

The main findings of this study are as follows: 1) Expression of TLR2 in HFSCs is decreased with aging and in a mouse model of obesity. 2) In young and healthy animals, TLR2 expression in HFs is cycle-dependent, with the highest expression in HFSCs during the initiation of the anagen phase. 3) The absence of TLR2 in HFSCs prolongs the resting phase of the hair cycle and significantly delays hair regeneration after injury. 4) TLR2 regulates the hair cycle primarily by inhibiting BMP signaling in HFSCs. 5) HFs continuously produce a metabolite of PUFAs, which acts as an endogenous TLR2 ligand and promotes hair growth through TLR2 activation in HFSCs. Besides reduced TLR2, aging is linked to low levels of its ligand in HFs. The stimulatory role of TLR2 signaling in HFs was demonstrated through both animal models and established human cell lines.

The lack of TLR2 appears to shift the balance between activating and inhibitory cues, leading to a resting phase that is approximately three times longer. This is a substantial impact considering that there are only 4-5 hair cycles in a mouse lifetime (Choi et al., 2021). The decrease in both TLR2 and its ligand observed in aging and associated conditions will inevitably impede the cyclic regeneration of hair follicles.

The immune system was shown to play a role in the activation of HFSCs, even in the absence of inflammation (Ali et al., 2017, Castellana et al., 2014, Pinho and Frenette, 2019). TLR2 expression increases at the onset of anagen when the immune response is reduced (Paus et al., 2003) and the hair follicle is most susceptible to pathogens. The upregulation of TLR2 in anagen may initially have a protective function. However, high TLR2 expression in undifferentiated vs. differentiated cells underscores its role in stem cell biology. TLRs, including TLR2, have been shown to play a critical role in stem cell functions in various organs (Lathia et al., 2008, Tomchuck et al., 2008, Trowbridge and Starczynowski, 2021). TLRs’ ligation and signaling can alter stem cell differentiation patterns (Collins et al., 2021, Nagai et al., 2006, Trowbridge and Starczynowski, 2021). Proinflammatory signaling can also activate HFSC proliferation, e.g., during injury (Wang et al., 2017, Chen et al., 2015). During the immune-privileged anagen phase of the hair cycle, TLR2 signaling may act as a key intrinsic factor in triggering HFSC activation.

The role of innate immunity in stem cell activation has mainly been linked to TLR3, shown to induce pluripotency in somatic cells through nuclear reprogramming (Lee et al., 2012, Sayed et al., 2015) and drive hair follicle (HF) neogenesis after tissue damage (Nelson et al., 2015). In contrast, we show that TLR2 drives a rapid inflammatory response and regulates the normal hair cycle and HF regeneration/neogenesis in injury.

TLR2 promotes the hair cycle by inhibiting the BMP pathway, a key regulator of hair follicle stem cell (HFSC) quiescence (Plikus et al., 2008, Hsu et al., 2011, Kandyba et al., 2013). Our study demonstrates that reducing excessive BMP signaling reactivates TLR2-deficient HFSCs, revealing a novel link between TLR2, BMP signaling, and the hair cycle. The only known instance of immune system-mediated BMP pathway inhibition occurs during the apoptosis of bulge-associated macrophages (Castellana et al., 2014, Wang et al., 2017).

The role of TLR2 in hair follicle regeneration in both normal hair growth and wound healing emphasizes the importance of understanding the nature of TLR2 ligands mediating these responses. At the site of injury, TLR2 can be activated by pathogens or by endogenously produced ligands, such as the oxidative product of PUFA, CEP, generated in abundance during wound healing (West et al., 2010, Yakubenko et al., 2018). CEP and TLR2 are both essential for hair regeneration, and their deficiency observed in pathologies such as aging and obesity might substantially impair hair growth. Exogenous application of CEP accelerates both wound closure (West et al., 2010) and HF regeneration through TLR2. In addition, both TLR2 and the application of CEP diminish inhibitory BMP signaling, suggesting that CEP and other TLR2 ligands could have therapeutic value for the treatment of hair loss related to burns, traumas, and other pathologies. It is intriguing that while CEP is almost exclusively generated at sites of injury and inflammation (Yakubenko et al., 2018, Xiong et al., 2022a), HFs continuously produce it, most likely by the means of MPO, an anti-bacterial enzyme, capable of generating CEP (Klebanoff et al., 2013).

Contrary to the trend observed in other tissues, where the accumulation of oxidation-generated CEP increases with aging (West et al., 2010), hair follicles (HFs) seem to exhibit depletion of CEP with aging. This decline in CEP levels might contribute to the reduced activity of hair follicle stem cells (Fuchs and Blau, 2020).

The role of CEP in TLR2-dependent HF growth and regeneration highlights the connection between oxidative stress and regenerative processes. Sustained reactive oxygen species (ROS) play a crucial role in proper regeneration, as seen in the tail amputation of Xenopus tadpole (Love et al., 2013). ROS enhance the differentiation of hematopoietic progenitors in Drosophila (Owusu-Ansah and Banerjee, 2009) and sustain self-renewal in neural stem cells (Le Belle et al., 2011). Additionally, ROS production in the skin has been linked to the activation of hair follicle stem cells (HFSCs) (Carrasco et al., 2015). We show the underlying mechanism for these observations, where oxidation-generated CEP triggers TLR2 activation, decreases inhibitory BMP signaling, and stimulates hair follicle growth and regeneration. TLR2 appears to serve as a common link between oxidative stress and tissue regeneration.

To summarize, our study highlights a novel role of TLR2 in promoting tissue regeneration during normal hair growth and wound healing. The identification of an endogenous TLR2 ligand produced by hair follicles presents a potential target for augmenting hair regeneration in the context of injury and aging, opening up new avenues for regenerative medicine.

Supporting information

eLife Assessment

In this manuscript by Xiong L et al., through TLR2, the authors have uncovered an important link between innate immune signaling and hair regeneration. The authors provide convincing evidence supporting the critical roles of TLR2 in sensing CEP levels in hair follicles, counteracting the action of BMP signaling, and facilitating the activation of HFSCs during the hair cycle and wound repair. Importantly, the authors also propose that decreased CEP production and TLR2 expression might be factors contributing to the decreased hair regeneration associated with age-related and obesity-related hair thinning and hair loss phenotypes.

Joint Public Review:

This paper's strengths are the interesting analysis of TLR signaling in hair follicle stem cell activation and the striking phenotype of the TLR2 cKO mice (but note below). The functional interrogation parts using HFSC-specific TLR2 genetic deletion are solid, and an endogenous regulator, CEP, is identified. The experiments reported in this manuscript are well-designed and presented. The authors provided extensive evidence supporting the roles of TLR2 signaling in regulating hair follicle stem cell functions. Importantly, the findings from this paper may have sustained impacts on our understanding of the roles of innate immunity in regulating tissue regeneration in the absence of inflammation.

The main evidence for the mechanistic analysis is based on fluorescence using immunohistochemistry, and here the expression analysis is not convincing. In addition, additional assays beyond immunolandscaping are needed to confirm the findings. The reviewers felt that your data substantiating the mechanism of interaction between TLR2 and BMP pathway needs bolstering.

Author Response

The following is the authors’ response to the original reviews.

To the reviewers.

We appreciate a detailed and deep review of our manuscript. Below are our comments and responses. Many requested data are present in the Supplementary figures of the manuscript. There seem to be two main concerns: one regarding the evidence of TLT2 expression in HFSCs; and second, regarding CEP/TLR2. As detailed below, we utilized 3 different methods to document TLR2 expression: TLR2-reporter mouse, staining for TLR2 and qPCR of isolated cells for TLR2. The source (the data are in Supplementary Fig. 5A, B and in references below) and nature of CEP (it is not a protein, but metabolic product of Polyunsaturated acid DHA oxidation by MPO amongst other ROS sources) are also explained below.

1. “The expression analysis of TLR2 is questionable. Many of the conclusions about the level of target genes are based on quantifying fluorescence intensity in microscopy images (e.g., TLR2 level in young or aged mice, BMP7 levels in mice with/without TLR2 KO). This could be strengthened by using qPCR to measure gene expression levels in FACS-sorted HFSCs, which would provide more accurate quantification. Additionally, the authors should test if the TLR2 antibody used is valid.”

In most instances we have used TLR2 reporter mouse, which presents an advantage over immunostaining. Fig.2 (A-H) shows expression of TLR2 reporter, not the staining with TLR2 abs. For selected experiments we utilized immunostaining with anti- TLR2 (Santa Cruz Biotechnology, sc-21759) antibody, which has been validated in our previous publication (see Michael G. McCoy and all. Endothelial TLR2 promotes proangiogenic immune cell recruitment and tumor angiogenesis. // Sci Signal. 2021 Jan 19; 14(666): eabc5371/doi: 10.1126/ scisignal.abc5371). In Fig.S2E of that manuscript we validated these abs using a knockout of TLR2. In the current paper, we further validate anti-TLR2 abs by showing its co-localization with the TLR2-GFP reporter (Fig. S1A).

We then confirmed reporter and immunostaining data by qPCR showing Tlr2 expression in FACS-purified mouse HFSCs in anagen, telogen, and catagen (Fig.2J), in mouse epidermal cells and FACS-purified HFSCs (Fig.2K), and FACS-purified HFSCs isolated from Control and TLR2HFSC-KO mice (Fig.4E).

As for the mechanistic link between TLR2 and BMP signaling was identified using RNAseq on FACS-purified HFSCs (supplementary Fig.4), then verified using qPCR (Fig.4E shows Bmp7,Bmp2, Bmpr1a ) and only then immunohistochemistry staining for BMP7 and phosphoSMAD1/5/9 was used (Fig.4A-D, F-H). Note that the large body of requested evidence is presented in Supplementary data. Other mechanistic links shown using qPCR include Nfkb2, Il1b, Il6, and Bmp7 in FACS-purified mouse HFSCs treated with BSA control or CEP (Fig.6Q,6R).

“As the reviewers note, it is not clear whether the TLR2+ signal is located at the basal side of bulge stem cells, basement membrane underlying bulge stem cells, or dermal sheath cells encapsulating bulge structure. Co-staining with basement membrane markers such as collagen and laminin or HFSC basal side membrane markers such as Itga6, Itgb1, and Itgb4 will clarify this. In addition, showing the expression pattern of TLR2 in full skin including epidermis and dermis would be helpful. As TLR2 is highly expressed in immune cells or blood endothelial cells, if the antibody staining is valid, strong positive signals should present in the cells. Moreover, testing the TLR2 antibody in Tlr2 knock-out mouse tissues would be an appropriate control experiment.”

Once again, in most instances we have used not the staining for TLR2 but TLP2 reporter mouse (Fig.2 legend). Anti-TLR2 abs have been verified in TLR2 KO as described above. Fig.2K shows comparison of Tlr2 mRNA expression in mouse epidermal cells to FACS-purified HFSCs by qPCR.

TLR2 signal is detected in several cell types within the hair follicle as well as in dermal cells surrounding the hair follicles, such as lymphocytes, resident tissue macrophages, fibroblast, and fibroblast precursors, etc. (https://www.proteinatlas.org/ENSG00000137462-TLR2/single+cell+type). In Author response image 1 below, white arrows point to the TLR2-positive cells around the hair follicle. In our paper, we focus on HFSC TLR2 and use the respective inducible tissue specific TLR2 KO. The contribution of TLR2 on other cell types can be assessed by the comparison of the phenotypes of global TLR2 KO, TLR2 KO-WT bone marrow chimeras and HFSC-specific TLR2 KO. The results are presented in both, main and supplementary figures (Fig.5D-I and SFig.5I-K shows global TLR2 KO, Fig.6H-I, SFig.5G-h shows bone marrow chimeras and Figs.3,4, 5 (J-M), Fig.5 (J-N) shows the main focus, HFSC-TLR2 KO. Overall, the phenotype (delay of hair regeneration after wounding) seems to be the strongest in TLR2 KO, whereas bone marrow chimeras and HFSCs phenotypes are comparable. Thus, TLR2 on bone marrow derived cells complements the main role for TLR2 on HFSCs.

Staining for TRLR2 (white), DAPI (blue) and Keratin 17 (purple) is shown

“The increase in expression of TLR2 during the hair follicle stem cell activation should be documented by FACS and/or qPCR. This is important because as noted by one of the reviewers.”

While original observation was done using both, a TLR2 reporter mouse and immunostaining, the data were confirmed by qPCR showing Tlr2 mRNA expression in FACS-purified mouse HFSCs in anagen, telogen, and catagen (Fig.2J).

“In Fig 1D, the authors mentioned that they re-analyzed published RNA-seq data (Greco et al., 2009) to show the increase of Tlr2 and Tlr6 expression in late telogen compared to early telogen. However, there is no RNA-seq data in that paper, but only microarray data of bulge vs HG comparison and dermal papillae cells (DP) in early, mid, late Telo. If the authors used DP data to show the increase of Tlr2 transcripts in late Telo, the analysis is completely wrong and has to be corrected. The problem is compounded by the fact that in other published HFSC RNA-seq datasets (Yang et al., Cell, 2017, Adam et al., Nature Cell Biology, 2020), the expression levels of Tlr2 and Tlr6 are very low (below 5 TPM). In Fig 1G, the authors also re-analyzed Morinaga et al., 2021 data to show the reduction of Tlr2 expression in HFSCs in high-fat diet mice. However, in the raw data of Morinaga et al., 2021 (GSE169173), Tlr2 expression FPKM values are below 1 in both normal diet and high-fat diet samples, which are too low to perform comparative analysis and are not statistically meaningful. Like Tlr2, the expressions of Tlr1 and Tlr6, which form heterodimer with TLR2, are almost 0. Thus, the authors should revisit the dataset and revise their analysis and conclusion.”

To document the existence of Tlr2 and Tlr6 expression in HFSCs, the authors should perform RNR-seq-based gene expression analysis by themselves. Otherwise, the authors' TLR2 expression analyses in Fig 1 are not convincing. These are serious issues that the authors will want to rectify so that eLIFE readers will not discount their findings and importance.”

It is correct, we analyzed a published array, not RNAseq data (Greco et al., 2009) using GEO2R tool which allowed us to compare the mRNA expression levels between early, middle, and late telogen in bulge CD34 positive cells. We changed the “RNA-seq” (the term was used incorrectly) to “RNA microarray” in the main text.

In our manuscript, TLR2 expression is documented not only in Fig.1, but also in Fig.2 and S.Fig.1. We utilized 3 different methods to document TLR2 expression: TLR2-reporter mouse, staining for TLR2 and qPCR of isolated cells for TLR2. Fig.2K shows comparison of Tlr2 mRNA expression in mouse epidermal cells to FACS-purified HFSCs by qPCR to document increased TLR2 expression on HFSCs. Likewise, Fig.2J shows qPCR for TLR2 on HFSC during various phases of hair growth.

“In Fig 2, to support the expression of Tlr2 in HFSCs, the authors utilized TLR2-GFP mice and showed the strong GFP expression in HFSCs, hair bulb, and ORS. However, as the expression data in Fig 1 are questionable, the GFP reporter data should be carefully analyzed with proper control experiments. For example, although TLRs are highly expressed in immune cells and endothelial cells, which are abundantly present in skin, Fig 2 data did show the GFP expression in these cells. Instead, the GFP signals looked very specific to epithelial compartments, which is odd. Again, to convince readers, the authors should provide more comprehensive analyses of expression patterns of TLR2-GFP mice in skin. Also, if the TLR2-GFP signals faithfully reflect the actual expression of Tlr2 mRNA, the GFP signals should increase in late telogen compared to early telogen. The authors should check whether TLR2-GFP expression follows this pattern.”

The specificity of TLR reporter was characterized in Price et al. , 2018. A Map of Toll-like Receptor Expression in the Intestinal Epithelium Reveals Distinct Spatial, Cell Type-Specific, and Temporal Patterns. Immunity, 49. Thus, TLR2 reporter mouse is well characterized (https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6152941/) and represents one of the best available tools to show TLR2 expression.

Expression of TLR2 on endothelial cells and validation of anti-TLR2 abs was performed in McCoy et al, Science Signaling as mentioned above. Also as discussed above we show a strong correlation between TLR2-GFP reporter expression and TLR2 expression using coimmunostaining with GFP and TLR2 antibodies with appropriate isotype-match non-immune antibodies as negative controls.

There is no doubt that TLR2 is expressed on immune, endothelial and epithelial cells. According to the Human Protein Atlas, TLR2 expression is identified in skin fibroblasts, keratinocytes, melanocytes, etc., so our findings are well supported by the literature (https://www.proteinatlas.org/ENSG00000137462-TLR2/single+cell+type). Indeed, we detected TLR2 in cells surrounding the hair follicle (see the pictures above). TLR2 signal was detected in nearly all niches of hair follicles including the CD34-positive cells.

In Fig.S1 we demonstrated an increased level of TLR2 in the late (competent) telogen compared to the early (refractory) telogen using immunostaining for TLR2-GFP. The results mirrored published RNA-array data in Fig.1D. Again, reporter and immunostaining results have been validated by qPCR for TLR2.

The levels of TLR2 might be heavily influences by the environment, i.e. pathogens availability. In this regard, note that mice for this study were kept in normal, not pathogen-free conditions.

“Overall, the existence of Tlr2 expression in HFSCs is still questionable. Without resolving these, genetic deletion of Tlr2 in HFSCs cannot be rationalized.”

In our manuscript, TLR2 expression is documented not only in Fig.1, but also in Fig.2 and S.Fig.1. We utilized 3 different methods to document TLR2 expression: TLR2-reporter mouse, staining for TLR2 and qPCR of isolated cells for TLR2. Besides these data, we show the functional responses to canonical TLR2 ligand, PAM3CSK4, and previously characterized endogenous ligand, CEP, using proliferation, western blotting and many other approaches. In numerous immunostainings we show co-localization of TLR2 and CD34 (Fig.2) using IMARIS surface rendering and colocalization tools. Our conclusions are further supported by published results as discussed above.

1. “The central conclusion of this study is that the activation of TLR2 can suppress BMP signaling; however, the molecular link between TLR2 and BMP signaling is still missing. Given the importance of this finding, it would be intriguing to further investigate how TLR2 activation suppresses BMP signaling. A better characterization of the molecular-level interaction between TLR2 and BMP signaling can further enhance the impact of this study.

-The published dataset should be re-analyzed, as some images and their quantification do not appear to be matched. Representative images should be used.”“In Fig 4, the authors propose that the activation of TLR2 pathway inhibits the BMP signaling pathway, which makes HFSCs quiescent. In TLR2-HFSC-KO, the authors showed that BMP7 is increased and pSMAD1/5/9 is sustained. The increase in BMP7 expression and SMAD activation should be demonstrated by additional assays. Are SMAD target genes activated in the cKO mice?”

This mechanistic link between TLR2 and BMP was originally identified by RNAseq, confirmed by qPCR and then by immunostaining for both, BMP7 and BMP pathway activation based on phosphoSMAD1/5/9 levels. The connection to BMP pathway was also shown by western blotting (S.Fig.4B,C). The rescue experiments have been performed using Noggin injections. According to our data, numerous SMAD target genes are upregulated in TLR2-HFSC-KO, such as Kank2, Ptk2b, Scarf2, Camk1, Dpysl2, as well as BMP2 and BMP7, and these changes were confirmed by qPCR analysis in Fig.4E. Additional evidence is shown in Fig.6, which demonstrates that endogenous TLR2 ligand, CEP-carboxyethylpyrrole, acts by a similar, BMP-dependent pathway. Also, Supplemental Fig.4 adds more details to this link. SFig.4B,C shows that TLR2 activation by canonical ligand PAM3CSK4 inhibits pSMAD levels induced by BMP (western blot is shown). At the same time, as anticipated PAM3CSK4 upregulated NFkB, however, little of no effect of BMP stimulation on NFkB is observed. To summarize: TLR2 affects both, BMP7 production and BMP induced downstream signaling judged by PhosphoSMADs. The later connection appears to go in one direction: TLR2 signaling affects BMP-induced pSMADs, however, BMP signaling does not seem to substantially change TLR2-dependent NFkB. We plan to delve into the intersection of these important pathways in future.

“Functionally, downregulation of BMP signaling by injecting Noggin, a BMP antagonist, in TLR2HFSC-KO mice induces HFSC proliferation. These functional data are solid. However, it is still curious how TLR2 signaling interact with BMP pathway molecularly. Is it transcriptional regulation or translational regulation? Perhaps, RNA-seq analysis of TLR2HFSC-KO could give some hints to answer this question. Furthermore, checking out other signaling pathways such as WNT/LEF1 and pCREB, which are important for hair cycle activation and NFkB, a downstream effector of TLR signaling would be helpful to interrogate mechanistic insights.”

As discussed above, TLR2 affects both, BMP7 production and BMP-induced downstream signaling judged by PhosphoSMADs. The later connection appears to go in one direction: TLR2 signaling affects BMP-induced pSMADs, however, BMP signaling does not seem to substantially change TLR2-dependent NFkB.

Indeed, in addition to BMP signaling, the Wnt signaling and β-catenin stabilization within HFSCs, known to trigger their activation (Deschene et al., 2014). However, this axis remained unchanged upon TLR2HFSC-KO (as shown in Supplementary Fig. 4J). There were several published reports on the crosstalk between TLR and BMP signaling such as (doi: 10.1089/scd.2013.0345. Epub 2013 Nov 7) showing that activation of TLR4 inhibits BMP-induced pSMAD1/5/8 and this connection requires NFkB. We probed NfkB activation, please, see the responses above.

However, we were not able to detect substantial effect of NFkB inhibition on BMP signaling in hair follicles (not shown).

1. “The function of CEP, a proposed endogenous ligand of TLR2, is still not clear. The authors imply that the decreased CEP level in aged mice could lead to deficient TLR2 signaling, which could further cause aging-associated hair regeneration defects. But this has not been demonstrated. What are the BMPs and pSmad1/5 levels in aged skin? Another important experiment to confirm the importance of this link during aging would be to inject CEP into the aged skin and examine whether this could restore hair regeneration in aged mice. Does CEP activate hair cycling during the endogenous pathway? What might be the source of CEP? Does CEP treatment activate BMP7 signaling? The authors should clarify these issues. The authors suggested that CEP is an endogenous ligand of TLR2, and administration of CEP induces hair cycle entry in a TLR2dependent manner. How potent is CEP in terms of HFSC activation? In Fig 6Q, CEP increases the expression of Nfkb2, Il1b, and Il6, but the fold changes are marginal. Also, if CEP is a critical ligand, the loss of CEP by a genetic deletion or a pharmacological inhibition should result in the delay of hair cycle entry. Furthermore, the source of CEP expression is curious. Is it expressed by HFSCs or dermal fibroblast or immune cells? Finally, comparing the effect of CEP to the effect of other bacterial origin Tlr2 ligands such as heat killed bacteria, purified microbial cell-wall components, and synthetic agonists (Pam3CSK4) would be helpful. It is curious if HFSC directly senses the bacterial materials and triggers hair follicle regeneration or are indirectly directed by immune cells and endothelial cells, which could be primary sensor.”

CEP is not a protein, it is an oxidative stress-generated metabolite of polyunsaturated fatty acid, DHA (https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5360178/), thus, it is impossible to generate a knockout of this molecule. As demonstrated in previous publications (https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2990914/, https://pubmed.ncbi.nlm.nih.gov/34871763/) CEP serves as a critical endogenous ligand supporting TLR2 signaling in the absence of pathogens. While other TLR2 endogenous ligands, such as HMGBs or HSPs exist (https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4373479/), CEP binds to TLR2 directly, and its generation is aided by MPO (myeloperoxidase) amongst other peroxidases and sources of reactive oxygen/nitrogen species. MPO (produced by immune cells amongst others) serves as an innate immunity response against pathogens, but it also generates CEP adducts (https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6034644/) adducts in both protein and lipid form. The knockout of MPO diminishes CEP generation in skin (PMC6034644), thereby demonstrating the causative relationship between CEP and MPO.

Additional immunostaining of mouse skin for Keratin 17 (purple), CEP (green) and MPO (red).

Similar staining is in S.Fig.5A and quantification is in S.Fig.5B.

Also, the above-mentioned manuscripts show that CEP effects are milder but overall comparable with canonical TLR2 agonists, PAM3SCK4. As we mention in the present manuscript, normal young mice’s tissues are devoid of CEP (which is generated in response to inflammation) with an exception of hair follicles. This is likely attributed to the secretion of MPO by hair follicles (PMID: 36402231) especially in conditions of inflammation (PMID: 32893875). Supplementary Fig.5A,B show that MPO is present at the high level in sebaceous gland (as a part of anti-microbial mechanism). Again, MPO is a secreted enzyme and it is likely to be a source of continuous DHA oxidation into CEP in hair follicles. We also document that both, TLR2 and CEP levels in hair follicles (but not in other tissues-an important point for CEP) are reduced in aging. Likewise, SFig.5A,B shows that MPO secretion in hair follicle is reduced by more than 60% in aging mice. Thus, it is likely that reduced MPO levels in aging hair follicle produce less CEP. Together with reduced TLR2 levels, the lack of CEP might contribute to hair loss in aging.

We show that similar to TLR2, CEP in hair follicles operates via a BMP-7 dependent pathway (see Fig.6). We also provide results using canonical bacterial ligand for TLR2, PAM3CSK4 whose effect on HFSCs proliferation is similar to CEP in a TLR2-dependent manner. TLR2 blocking approaches were used (Supp. Fig.4B, C, D, E, Supp. Fig.5D-5F). It remains to be seen whether CEP is required for the normal hair cycling and whether its administration might improve hair loss in aging subjects.

“The impacts of CEP/TLR2 on proliferation of keratinocytes is still weak. How much of this effect is a result of NFkB activation, and how much is simply due to inhibiting BMP signaling?

Impact of TLR2 on proliferation was demonstrated using a variety of mouse models, from global TLR2 KO to bone marrow chimeras to HFSCs-specific TLR2 KO, again using multiple approaches. The same applies to the effects of CEP as well as to canonical TLR2 ligand, PAM3CSK4, which were demonstrated both in vivo and in culture to be TLR2-dependent (Fig.6MO) and Supplementary Fig.4E-D). As for NFkB connection, see our responses above. It seems that the connection between TLR2 and BMP pathway occurs independently of NFkB activation.

1. The links between TLR2 pathway and aging and obesity are only correlative. Although the authors suggest that the reduction of TLR2 expression in aging and obesity may diminish hair growth (Fig 1), there is no direct functional evidence that supports this possibility. If the authors wish to make this claim, they should test the roles of TLR2 and CEP in aging and obesity conditions.”

We show that both, TLR2 and CEP are reduced in aging, and that this pathway contributes to hair cycling and regeneration upon wounding, we do not wish to claim more.

1. More minor points:

“Fig.4: The Noggin treatment in TLR2 KO mice is an important experiment. However, it is unclear why Noggin only enhances proliferation (Ki67 level) in HG but not in the bulge. This discrepancy should be addressed.”

As we showed in Fig. 3B-3F, TLR2 HFSC-KO mice have prolonged first telogen. Noggin treatment at the first postnatal telogen promotes telogen to anagen transition in TLR2HFSC-KO characterized by the activation of HG cells prior to the bulge cells. According to the literature, the bulge cells remained silent during the late telogen, however, HGs became Ki67- positive and the proliferation of HG cells contributed to the telogen-to-anagen transition.

(https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2668200/

https://www.sciencedirect.com/science/article/pii/S0022202X15404518?via%3Dihub

https://journals.biologists.com/jcs/article/114/19/3419/34892/Hair-follicle-predetermination).

“Fig.5: Does TLR2 cKO slow down wound healing, in addition to affecting pigmentation and the number of hair follicles?”

In our previous publication, we demonstrated that deletion of TLR2 in HFSC does not affect wound healing process. Instead, endothelial TLR2 promotes wound vascularization and healing.

(see Xiong and all. Timely Wound Healing Is Dependent on Endothelial but Not on Hair Follicle Stem Cell Toll-Like Receptor 2 Signaling.// Journal of Investigative Dermatology, Volume 142, Issue 11, November 2022, Pages 3082-3092.e1).

“There is no panel B in Fig.4. There is no image in Fig 4D. Please correct this properly.”

We corrected Fig.4

“Discussion: The constant production of CEP in homeostatic skin and in the absence of inflammation should be further discussed. Additionally, the possible causes of reducing CEP levels during aging should also be further discussed.”

We explained the sources of CEP generation, such as MPO as a one of the key enzyme, above.

The data on MPO levels in hair follicles of young and old mice are presented in Supplementary Fig.5A,B. Since we previously shown that MPO produces CEP from DHA (PMC6034644), the reduction in MPO in aging is likely to contribute to reduced CEP levels.

Version 2

Abstract

Abstract

The etiology of hair loss remains enigmatic, and current remedies remain inadequate. Transcriptome analysis of aging hair follicles uncovered changes in immune pathways, including Toll-like receptors (TLRs). Our findings demonstrate that the maintenance of hair follicle homeostasis and the regeneration capacity after damage depends on TLR2 in hair follicle stem cells (HFSCs). In healthy hair follicles, TLR2 is expressed in a cycle-dependent manner and governs HFSCs activation by countering inhibitory BMP signaling. Hair follicles in aging and obesity exhibit a decrease in both TLR2 and its endogenous ligand carboxyethylpyrrole (CEP), a metabolite of polyunsaturated fatty acids. Administration of CEP stimulates hair regeneration through a TLR2-dependent mechanism. These results establish a novel connection between TLR2-mediated innate immunity and HFSC activation, which is pivotal to hair follicle health and the prevention of hair loss and provide new avenues for therapeutic intervention.

Introduction

Hair follicles (HFs), represent one of the best examples of mini-organs with the ability to regenerate throughout life, which, in turn, relies on the proliferation and differentiation of HF stem cells (HFSCs) within hair bulge (Fuchs and Blau, 2020, Sakamoto et al., 2021). The cyclic renewal of HFs is orchestrated by the interplay between inhibitory and stimulatory signals (Plikus et al., 2011). Despite the immune privileged status of HFs, they have a unique microbiome and immune system, including resident macrophages and other immune cells (Fuchs and Blau, 2020, Bertolini et al., 2020, Paus et al., 2003). Components of the HF immune system have been implicated in regulating the HF cycle and its regeneration (Rahmani et al., 2020, Di Domizio et al., 2020). Given their exposure to pathogens, HFs are equipped with innate immune receptors, particularly Toll-like receptors (TLRs), which detect and respond to pathogens by stimulating the secretion of defensins (Di Domizio et al., 2020, Selleri et al., 2007).

Toll-like receptors (TLRs) play a key role in recognizing and responding to either pathogen- or damage-associated molecular patterns, mediating the cytokine response. However, the role of TLRs extends beyond this function, as they have been shown to directly promote tissue regeneration and homeostasis in multiple tissues, particularly in stem and progenitor cells. TLRs regulate hematopoietic and intestinal stem cell renewal, proliferation, and apoptosis (Nagai et al., 2006, Tomchuck et al., 2008). The role of innate immune responses in the tissue-healing benefits of stem cell therapy has been clearly demonstrated (Vagnozzi et al., 2020). Moreover, TLR activation is a critical component of the reprogramming or transdifferentiation of adult cells into pluripotency (Lee et al., 2012), emphasizing the close coordination between innate immunity, cell transformation, and regeneration.

Multiple reports connect altered HFs immunity to hair loss, including a breakdown of immune privilege in alopecia areata (Rahmani et al., 2020). Likewise, androgen, which is tightly linked to TLR activation, was shown to influence the innate immunity of HFs in androgenic alopecia (Sawaya, 2012). The decline of innate immunity processes due to aging or conditions like obesity is widely recognized and these conditions are causatively associated with hair thinning and loss. (Andersen et al., 2016, Shaw et al., 2013, Palmer and Kirkland, 2016, Ghanemi et al., 2020). Alopecia patients often have higher body weight index and weight compared to healthy individuals (Bakry et al., 2014). Increased body weight index is linked to more significant hair loss severity in adults (Goette and Odom, 1976) and a higher prevalence of hair disorders in children and adolescents (Mirmirani and Carpenter, 2014). Mouse models support these findings, showing that activation of innate immunity through pathogen signals might lead to alopecia (Shin et al., 2018) and that high-fat diets inducing obesity cause hair thinning through HFSC depletion (Morinaga et al., 2021).

Our previous studies have shown that activating endothelial TLR2 by endogenous ligands such as CEP (a product of PUFA oxidation) promotes wound healing and tumor angiogenesis (West et al., 2010). Deletion of TLR2 from endothelial cells reduces tumor size by diminishing its vasculature (McCoy et al., 2021). In wounded skin, endothelial TLR2 is crucial for tissue regeneration through increased proangiogenic cytokine secretion (Xiong et al., 2022b). Although PUFAs have been shown to benefit hair growth by extending the anagen phase and promoting cell proliferation and hair shaft elongation (Munkhbayar et al., 2016), the role of innate immunity and in particular, TLR2 in the hair follicle cycle remains unknown.

Using animal models and human cell lines, we show a new function of TLR2 in the hair follicle cycle in homeostasis and hair follicle regeneration in injury. Furthermore, we demonstrate that an endogenously produced PUFA metabolite CEP serves as a TLR2 ligand in the hair bulge, promoting hair regeneration and growth through TLR2. In conditions associated with hair loss, i.e. aging and obesity, both TLR2 and its ligand are substantially depleted in HFs.

Results

TLR2 in HF declines due to aging and obesity

To assess whether and how aging affects hair follicle innate immunity, we analyzed available RNA sequencing data of mouse HFSCs (Doles et al., 2012). Pathway analysis revealed that innate and adaptive immunity, as well as TLR signaling, were among the top dysregulated pathways (Fig. 1A). Notch, JAK-STAT, TGF-β, and Wnt, and other pathways essential for HF regeneration were also altered by aging (Fig. 1A). Notably, the level of Tlr2 mRNA in HFSCs of old mice was two-fold lower compared to young mice (Keyes et al., 2013). In addition, TLR2 at the protein level was substantially lower in 13 months old mice compared to 2 months old mice (Fig. 1B, 1C).

HFSCs downregulate TLR2 in response to stress like a high-fat diet and aging.

A. Dysregulated pathways in old vs young mouse HFSCs. The top pathways are labeled red.

B. Representative confocal images of telogen hair follicles from young and old mice immunostained for TLR2 and CD34 demonstrate decreased TLR2 intensity in HFSC (CD34-positive) of old mice. Scale bars are 50 μm. The middle and right panels show a magnified view of the boxed area. Scale bars are 20 μm.

C. Quantification of TLR2 fluorescent intensity in images from B showing significantly lower TLR2 expression in HFSCs from the old mice. N =6 for each group.

D. GEO2R analysis of published RNA data from sorted follicle populations in the 2nd telogen to anagen transition demonstrates the increased level of TLR2 mRNA accompanied by the activation of TLRs signaling downstream.

E. Representative confocal images showing TLR2 expression in hair follicles from mice fed with a normal diet (ND) or high-fat diet (HFD). CD34 is an HFSC marker. Scale bars are 50 μm. Magnified images demonstrate decreased TLR2 intensity in HFSC (CD34-positive) of mice after HFD. Scale bars are 20 μm.

F. Quantification of TLR2 fluorescent intensity in images from E showing significantly lower TLR2 expression in HFSCs from HFD-fed mice. N = 7 and 6 for ND and HFD groups respectively. AU, arbitrary unit.

G. TLR2 mRNA expression in HFSCs from mice fed with ND or HFD for 4 days or 3 months. Data regenerated from published RNA sequencing dataset GSE131958. N = 3 for each group.

All bar graphs are mean ± s.e.m. Non-parametric Mann-Whitney test (C, G) or unpaired two-tailed t-test (F) was used to determine statistical difference.

At the same time, the normal hair cycle is marked by an increase in TLR2 mRNA levels prior to HFSC activation during telogen, according to the analysis of existing RNA microarray data (Fig. 1D) (Greco et al., 2009). TLR2 mRNA levels are highest among other TLRs, with TLR1 and TLR4 mRNA showing the opposite pattern. TLR6 may act as a co-receptor for TLR2 as its expression pattern is similar. While the downstream TLR2 signaling molecules, IRAK1 and MYD88, are also upregulated, mirroring the TLR2, IRAK1 inhibitor TOLLIP is suppressed. Together, the results suggest the critical regulatory role of the entire TLR2/TLR6 pathway in the hair follicle cycle.

High-fat diet-induced obesity causes hair thinning and subsequent loss (Morinaga et al., 2021). In our model, a high-fat diet causes a nearly 2-fold decline in TLR2 protein level in HFSCs, compared to normal diet-fed mice (Fig. 1E, 1F). Further, RNA sequencing data reveal that 3 months of a high-fat diet is sufficient to reduce TLR2 levels in HFSCs by more than 2-fold compared to control mice (Fig. 1G) (Morinaga et al., 2021). Thus, our results and analyses of existing datasets demonstrate that conditions causatively associated with hair thinning and loss, such as aging and obesity, result in a dramatic depletion of TLR2 in HFSCs suggesting a possible regulatory role for TLR2 in hair follicles.

TLR2 is upregulated during HFSC activation

The expression of TLR2 during a normal hair cycle was assessed using a previously characterized TLR2-GFP reporter mouse (Fig. 2), one of the best tools for the analysis of TLRs in vivo (Price et al., 2018). The correlation between the reporter and TLR2 protein expression was confirmed (Supplementary fig. 1A, 1B). The hair follicle cycle was verified by H&E staining (Supplementary fig. 1C). During telogen, a dormant stage for HFSCs, TLR2 was found in the bulge, secondary hair germ (sHG), dermal papilla (DP), and outer root sheath (ORS) (Fig. 2A). During regenerative anagen, TLR2 expression was detected in the DP and all sHG-derived progenitor cells (Fig. 2B, 2C) and in quiescent bulge stem cells (Fig. 2D, 2E). All cells derived from bulge stem cells (ORS) and sHG (hair shaft and inner root sheath, IRS) were positive for TLR2 (Fig. 2C). In catagen, TLR2 was abundant in the new bulge and sHG formed from ORS cells (Hsu et al., 2011) (Fig. 2F, 2G). While TLR2 was expressed in early sHG lineage (including IRS) (Fig. 2C), it was absent in mature (Fig. 2D, 2E) and regressing (Fig. 2F) IRS. This shows that TLR2 is abundant in stem cells but declines upon differentiation. The second telogen’s old and new bulges expressed TLR2 (Fig. 2H). TLR2 was present in sHG and DP during the second telogen (Fig. 2H) and increased in the late (competent) telogen compared to the early (refractory) telogen (Supplementary fig. 1D, 1E). The highest level of TLR2 occurred during active anagen compared to quiescent telogen and catagen (Fig. 2I). Quantitative PCR revealed that the Tlr2 mRNA level in HFSCs was 5- and 2.3-fold higher in anagen than in telogen and catagen, respectively (Fig.2J). Notably, analysis of existing RNA sequencing data using FACS-sorted cells (Lorz et al., 2010) confirmed that TLR2 expression was significantly higher in HFSCs than in epidermal or non-stem cells (Fig. 2K). Thus, TLR2 is enriched in HFSCs, and its expression increases during activation.

TLR2 is enriched in HFSCs and is upregulated during HFSC activation.

TLR2-GFP reporter mouse skin sections were immunostained with anti-GFP to assess TLR2 expression in the hair follicles.

A. Representative confocal images of P21 first telogen hair follicle immunostained for TLR2-GFP, CD34 (bulge stem cells), P-cad (secondary hair germ (sHG)), and DAPI (nuclei). The green color in the surface rendering panel represents TLR2 expression, and other surfaces show co-localization between TLR2 and specific markers. TLR2 is present in bulge, sHG, and dermal papilla (DP) cells. P represents postnatal days. Scale bar is 10 μm.

B. TLR2-GFP in P28 anagen was co-immunostained with CD49f of basement membrane outlining the DP Scale bar is 10 μm.

C. TLR2 is co-localized to the sHG lineage (P-cad+ layers), DP, and outer root sheath (ORS) lineage. Scale bar is 20 μm.

D. TLR2-GFP in P28 anagen was co-immunostained with CD34 in old bulge (D) and Ker5 in ORS (E) revealing TLR2 localization to the old bulge, ORS, but not inner root sheath (IRS). Scale bars are 20 μm.

F. Co-immunostaining of TLR2-GFP in P38 catagen hair follicle with Ker5 in ORS lineage cells showing co-localization of TLR2 with ORS and bulge. Scale bar is 20 μm.

G. P41 late catagen hair follicle immunostained for TLR2 and CD34 showing co-localization of TLR2 to the old bulge, new bulge, sHG, and DP Scale bar is 20 μm.

H. P53 second telogen hair follicle immunostained for TLR2, CD34, and P-cad reveals co-localization of TLR2 to the bulge, sHG, and DP Scale bar is 20 μm.

I. Quantification of TLR2 fluorescent intensity in bulge cells at different phases showing TLR2 upregulation in anagen. N = 3 for each group.

J. QPCR analysis of Tlr2 mRNA expression in FACS-purified mouse HFSCs in anagen, telogen, and catagen. N = 3 or 4 per group.

K. QPCR analysis of Tlr2 mRNA expression in mouse epidermal cells and FACS-purified HFSCs showed significantly higher Tlr2 expression in HFSCs compared with raw epidermal cells. N = 6 mice per group.

All bar graphs are mean ± s.e.m. Two-tailed unpaired t-test (K) or Kruskal-Wallis test with Dunn’s post hoc test (I, J) was used to determine statistical difference.

Deletion of TLR2 in HFSCs delays anagen onset in the normal hair cycle

To address the role of TLR2 in the hair cycle, we generated an HFSC-specific inducible Tlr2 knockout mouse line (TLR2HFSC-KO) and deleted Tlr2 during the first postnatal telogen (Xiong et al., 2022b). In TLR2HFSC-KO mice, the telogen phase was substantially prolonged compared to control (TLR2lox/lox) mice, as summarized in the schematic (Fig. 3A). Melanogenesis and anagen onset are tightly coupled (Muller-Rover et al., 2001). Thus, the pink skin color at P21 marks the first postnatal telogen. The onset of anagen in control mice was indicated by the change in skin color to gray or black at P26 in control mice. TLR2HFSC-KO mice at this age did not enter anagen, as evidenced by the delayed darkening of their skin (Fig. 3A-3C). This was further confirmed by skin section analysis at P21, P26, and P35 (Fig. 3D). At P26, control mice displayed pigmented anagen hair follicles with enlarged bulbs located deeply in the hypodermis, while nearly all follicles of TLR2HFSC-KO mice were ∼5 fold shorter and remained in the dermis on top of adipose tissue, a characteristic of telogen. On day p35, we observe partial entrance into anagen in TLR2HFSC-KO skin while the skin color of TLR2HFSC-KO mice remains pink (Fig. 3D-3F). HFSCs were activated as early as at P24 in control mice based on positive Ki67 staining in sHG and bulge region (Fig. 3G), while most cells in TLR2HFSC-KO sHG (Fig. 3H) and bulge (Fig. 3I) remained quiescent. At P25, control mice exhibited a large cluster of P-cad+ cells encapsulating DP within the transformed sHG (Fig.3K), whereas the sHG of TLR2HFSC-KO mice remained small and inactive (Fig. 3J, 3K and Supplementary fig. 2). Despite the substantial delay in anagen onset, the morphology of hair follicles and expression of established HFSC markers, including Ker15, CD34, and Sox9, were normal in TLR2HFSC-KO mice (Supplementary fig. 3). Thus, TLR2 in HFSCs is essential for HFSC activation and progression of the hair cycle.

Deletion of TLR2 in hair follicle stem cells delays anagen onset.

A. Schematic of RU486-mediated Cre induction and dorsal skin pigmentation change (gradient bars) in TLR2lox/lox and TLR2HFSC-KO mice.

B. Representative images of shaved TLR2lox/lox and TLR2HFSC-KO mice showing different phases of the hair cycle. The TLR2lox/lox mouse transitions from telogen (pink skin) to anagen (grey/black skin) at P26 and a full hair coat is developed by P35. The TLR2HFSC-KO mouse exhibits a prolonged telogen (P21-P30-P35). Representative images of at least 10 mice in each group.

C. Bar graph showing the length of first postnatal telogen starting from P21 measured by skin color change from B. N = 10 per group.

D. Representative H&E staining of dorsal skin at indicated time points showing prolonged telogen in TLR2HFSC-KO mice. Scale bars are 50 μm.

E. The length of hair follicles at P26 from images in D. 50 hair follicles from 3 mice per group were used for quantification.

F. Percentages of telogen or anagen hair follicles at P26 from D. N = 4 mice per group.

G. Representative confocal images of P21 and P24 first telogen hair follicles from TLR2lox/lox and TLR2HFSC-KO mice immunostained for CD34, Ki67, and DAPI. Stars label the hair shaft. Scale bars are 20 μm.

H, I. Quantification of images in G shows a diminished number of Ki67+ cells in sHG (H) and in CD34+ bulge (I) in TLR2HFSC-KO mice compared to TLR2lox/lox at P24. N = 6 and 4 mice for TLR2lox/lox and TLR2HFSC-KO group respectively.

J. Representative confocal images of P21 and P25 dorsal skin sections from TLR2lox/lox and TLR2HFSC-KO mice immunostained for P-cad and DAPI showing changes in the size of sHG. Scale bars are 20 μm.

K. Quantification of sHG size in panel K shows enlarged sHG in TLR2lox/lox mice compared with TLR2HFSC-KO mice. N=4 mice for p 25 TLR2lox/lox, and N=5 mice for TLR2HFSC-KO. Statistical significance was determined using a non-parametric Mann-Whitney test. All data are mean ± s.e.m.

TLR2 regulates HFSC activation by interacting with BMP signaling pathway

The relationship between WNT and BMP signaling is central to the cyclic growth of hair follicles (Plikus et al., 2008). Anagen initiation is triggered by Wnt/β-catenin activation, while BMP signaling suppresses HFSC activation and its reduction is necessary for HFSC activation. Indeed, during the early (refractory) phase of the second telogen, HFSCs exhibit elevated BMP signaling as evidenced by high levels of BMP7 protein and pSMAD1/5/9, downstream targets of BMP signaling, compared to the late (competent) phase (Fig. 4A-4D). In contrast, BMP7 and its effectors, ID1 and ID2, are decreased during the late telogen based on our analysis of existing RNAseq datasets (Greco et al., 2009) (Supplementary Fig. 4A).

TLR2 interacts with BMP pathway to regulate the hair cycle.

A. Representative confocal images of BMP7 staining in hair follicles of dorsal skin in early (p46) and late (p69) second telogen. Scale bars are 10 μm.

B. Quantification of BMP7 fluorescent intensity from A showing diminished BMP7 expression during the second telogen from the early to late phases. N = 4 per group.

C. Representative confocal images of pSMAD1/5/9 staining in hair follicles of dorsal skin in early (p46) and late (p69) second telogen. Scale bars are 10 μm.

D. Quantification of pSMAD1/5/9+ positive cells in CD34+ bulge stem cells demonstrates a decrease of pSmad1/5/9 expression in late telogen. N = 4 per group.

E. QPCR analysis reveals dysregulation of BMP singling molecules in HFSCs lacking TLR2. N=4 mice for Control and BMP2, N=3 mice for BMP7 and BMPr1A.

F. Representative confocal images of BMP7 staining in hair follicles from TLR2lox/lox or TLR2HFSC-KO mice. Scale bars are 10 μm. Stars label hair shaft.

G. Quantification of BMP7 fluorescent intensity from F showing higher BMP7 expression in TLR2HFSC-KO mice. N = 4 per group.

H. P21 and P24 dorsal skin sections from TLR2lox/lox and TLR2HFSC-KO mice immunostained for CD34, pSmad1/5/9, and DAPI. Scale bars are 10 μm.

I. Quantification of pSmad1/5/9+ cells in CD34+ bulge stem cells in P24 dorsal skin from H. N = 4 and 5 for TLR2lox/lox and TLR2HFSC-KO respectively.

J. Representative confocal images of dorsal skin sections from TLR2HFSC-KO mice treated with BSA or noggin immunostained for CD34, pSmad1/5/9, and DAPI. Star labels the hair shaft. Scale bars are 10 μm.

K. Quantification of pSmad1/5/9+ cells in CD34+ bulge stem cells from images in J. N = 5 per group.

L. Immunostaining for Ki67 and DAPI in dorsal skin sections from TLR2HFSC-KO mice treated with BSA or noggin. Scale bars are 10 μm.

M. Quantification of images in L showing an increase in Ki67+ cells in sHG of noggin-treated compared to BSA-treated TLR2HFSC-KO dorsal skin. N = 5 per group.

N. Representative confocal images of Ki67 and DAPI immunostaining of dorsal skin sections from TLR2HFSC-KO mice treated with BSA or noggin. Arrows point to hair follicles with Ki67+ cells in the sHG. Scale bars are 20 μm.

O. Quantification of images in N showing percentages of hair follicles with Ki67+ cells in sHG. N = 5 per group.

P. BSA- or noggin-treated TLR2HFSC-KO mouse dorsal skin immunostained for P-cad and DAPI. The dashed line outlines the sHG. Scale bars are 10 μm.

Q. Bar graph showing significantly larger sHG in noggin-treated TLR2HFSC-KO mice. N = 5 per group.

Mann-Whitney test was used to determine the statistical significance. All data are mean ± s.e.m.

To assess the possible connection between the TLR2 signaling and BMP pathway in human cells, we activated TLR2 and BMP signaling in human epidermal keratinocytes (NHEK) using a canonical TLR2 agonist (Pam3CSK4) and BMP4, respectively. As anticipated, BMP4 promoted the phosphorylation of its downstream target SMAD1/5/9. However, simultaneous co-activation of TLR2 diminished BMP4 signaling (Supplementary fig. 4B, 4C).

Likewise, stimulation of human HFSC with canonical TLR2 agonist Pam3CSK4 promoted cell proliferation by 1.5-fold compared to controls. Notably, this effect was diminished in the presence of a TLR2-blocking antibody (Supplementary fig. 4D, 4E). These results reveal that TLR2 activation on human HFSC augments their proliferation.

To gain a deeper understanding of TLR2’s role in HFSC activation, we profiled the transcriptome of HFSCs lacking TLR2 expression (Supplementary fig. 4F). The results showed that TLR2 deletion dysregulated 486 genes, many of which were involved in both the hair cycle and innate immunity. The most affected pathways included innate immunity response and TLR2 signaling, with its downstream target NF-kappaB (Supplementary fig. 4G). This profile somewhat resembles changes observed in aging models (Fig. 1A).

BMP pathway is altered in TLR2 KO HFSCs

Since TLR2 suppresses BMP signaling and promotes HFSC proliferation, we assessed whether the delayed anagen in TLR2HFSC-KO mice might be associated with the BMP pathway. QPCR analysis reveals that several key components of the BMP pathway were dysregulated in HFSCs lacking TLR2 (Fig.4E). Among those, the most notable changes were observed for BMP7, which was upregulated by ∼4-fold in TLR2-null HFSCs compared to controls (Fig. 4E). This was substantiated by co-staining of tissue sections for BMP7 and CD34, which demonstrated a ∼2-fold increase in BMP7 on HFSCs of TLR2HFSC-KO mice as compared to the control (Fig. 4F, 4G). Activation of BMP signaling was assessed by pSMAD1/5/9 positive staining in HF. Quantification revealed a ∼15-fold higher level of pSMAD1/5/9 in TLR2-null mice (TLR2KO) as compared to controls (wild type) (Supplementary fig. 4H,4I). The significant increase in BMP signaling observed was attributed to the absence of TLR2 in hair follicle stem cells (HFSCs). This was evidenced by a comparable 14-fold increase in BMP signaling in follicles of TLR2HFSC-KO mice compared to control during the first postnatal telogen phase, thereby ensuring the preservation of follicles in the dormant telogen stage (as shown in Fig. 4H and 4I). Simultaneously, the Wnt signaling and β-catenin stabilization within HFSCs, known to trigger their activation (Deschene et al., 2014), remained unchanged between control and TLR2HFSC-KO mice (as shown in Supplementary Fig. 4J).

BMP antagonist rescues defects caused by the lack of HFSCs TLR2

To demonstrate that an altered BMP pathway is, indeed, responsible for the phenotype of TLR2 knockout in HFSCs, we utilized intradermal injection of noggin, a well-known inhibitor of BMP signaling (Botchkarev et al., 2001, Botchkarev et al., 1999), to block the upregulated BMP signaling in TLR2HFSC-KO mice. As a result, noggin injection diminished activation of BMP signaling by >10-fold in TLR2HFSC-KO mice as assessed by pSMAD1/5/9 staining of HF (Fig. 4J, 4K). Moreover, noggin promoted activation of TLR2HFSC-KO HFs while the HFs in BSA-treated TLR2HFSC-KO mice remained quiescent. Noggin treatment of TLR2HFSC-KO mice dramatically upregulated cell proliferation within sHG as evidenced by Ki67+ cells (Fig. 4L, 4M), promoting ∼2.5 fold increase in activated follicles (Fig. 4N, 4O), thereby contributing to nearly 2-fold larger sHG (Fig. 4P, 4Q) as compared to BSA-treated TLR2HFSC-KO mice. Thus, curbing suppressive BMP signaling in TLR2HFSC-KO mice can reactivate their HFs, demonstrating a causative connection between TLR2 and BMP pathways in the hair cycle.

HFSC TLR2 governs hair regeneration upon injury

High expression of TLR2 and its critical role in HFSC activation during the hair cycle prompted us to test the role of HFSCs TLR2 in an injury model where cells are more likely to be exposed to TLR2 ligands. First, we compared TLR2 levels in HFs in wounded and healthy skin using TLR2-GFP reporter mouse (Fig. 5A). In healthy skin, HFSCs upregulated TLR2 during their transition from middle to late telogen (day 5 to day 10) (Fig. 5B upper panels, and grey bars in fig. 5C), consistent with RNA sequencing results (Greco et al., 2009). This increase in TLR2 precedes HFSCs activation during the normal cycle. However, in wound HFSCs, TLR2 was upregulated immediately after an injury resulting in 1.5-fold higher expression compared to normal unwounded skin (Fig. 5B, 5C).

HFSC TLR2 is crucial for wound-induced hair follicle regeneration.

A. Schematic of wound healing assay using TLR2-GFP reporter mouse. Full-thickness wounds on the dorsal skin of TLR2-GFP mice were created. Normal unwounded skin and the skin adjacent to the wound were harvested at different time points.

B. Representative confocal images of normal and wounded skin from TLR2-GFP mice at different time points post-injury immunostained for TLR2-GFP and CD34. Scale bars are 20 μm.

C. Quantification of TLR2 fluorescent intensity per bulge cell in hair follicles from B shows increased TLR2 level in hair follicles from wounded skin as compared to normal skin. N=3 for Day 1 and Day 5, and N=4 for Day 10 per group.

D. Representative photographs showing hair regeneration on the dorsal skin and inner skin flaps at indicated time post-injury in WT and TLR2 global knockout (TLR2KO) mice. Diminished hair growth around the wound is apparent in TLR2KO skin from day 14 through 21 post-injury compared to WT skin. The inner skin flaps from TLR2KO at day 14 post-injury show an absence of pigmented hair bulbs and skin pigmentation. Scale bars are 5mm for dorsal skin and 1mm for inner skin flaps.

E. Quantification of pigmented dorsal skin area around the wound from images in D shows diminished pigmentation in TLR2KO skin compared with WT skin at all time points post-injury. N=4 per group.

F. Representative confocal images of skin adjacent to wound immunostained for Ki67 and DAPI. Scale bars are 50 μm.

G. Bar graph showing diminished Ki67 fluorescent intensity in the skin adjacent to wound in TLR2KO mouse compared to WT mouse from images in F. N = 4 and 7 for WT and TLR2KO respectively.

H. Representative confocal images of skin adjacent to wound immunostained for Ker17 and DAPI. Scale bars are 100 μm.

I. Quantification of hair follicle numbers from images in H reveals a significant decrease in regenerated hair follicles in TLR2KO skin compared with WT skin. N = 5 and 7 for WT and TLR2KO respectively.

J. Representative photographs showing a lack of hair regeneration and skin pigmentation around the wound on the dorsal skin of TLR2HFSC-KO mice compared with TLR2lox/lox mice on day 17, day 21, and day 24 post-injury. Scale bars are 2mm.

K. Quantification of pigmented skin area around the wound during 14-28 days post-injury showing significantly smaller pigmented skin area in TLR2HFSC-KO mice compared with TLR2lox/lox mice. N = 4 per group.

L. Representative confocal images of wounded skin from TLR2lox/lox and TLR2HFSC- KO mice stained for CD34 and pSmad1/5/9. Scale bars are 10 μm.

M. Quantification of images from L showing more pSmad1/5/9+ cells in TLR2HFSC-KO wounded skin. N = 3 per group.

Mann-Whitney test was used to determine the statistical significance. All data are mean ± s.e.m.

Hair regeneration after injury represents a substantial part of the healing process (Wang et al., 2017, Chen et al., 2015, Abbasi and Biernaskie, 2019). We next assessed the role of TLR2 using age- and gender-matched WT and TLR2KO mice. The lack of TLR2 visibly impaired hair regeneration after wound healing (Fig.5D). On day 14 post-injury, HFs in WT mice entered precocious anagen judged by a spot of pigmented skin, which, on day 21 developed into a black hair patch (Fig. 5D). In contrast, the follicles of TLR2KO mice remained quiescent lacking regenerated HFs around wounds even after 21 days post-injury (Fig. 5D, 5E). At this point, pigmented skin area in WT was ∼9-fold larger than in TLR2KO mice. Skin flaps showed substantial pigmentation and growing hair bulbs around wounds in WT, indicative of active anagen. In contrast, the TLR2KO skin flap was devoid of pigmentation, consistent with telogen (inner skin flap in fig. 5D). Ki67 staining confirmed an increase in HFs activation in WT but not in TLR2KO skin (Fig. 5F, 5G). The resulting density of regenerated HFs based on Ker17 staining in WT was 2-fold higher than in TLR2KO mice (Fig. 5H, 5I). Most importantly, this effect was dependent on TLR2, specifically on HFSCs, since TLR2HFSC-KO mice exhibited a similar phenotype with a dramatic reduction in pigmentation and hair growth compared to control mice (Fig. 5J, 5K). The upregulation of pSmad1/5/9 in TLR2HFSC-KO wounds compared to controls demonstrates that similar to the HF cycle scenario, increased BMP signaling might contribute to diminished HF regeneration (Fig. 5L, 5M). Thus, the TLR2-BMP axis in HFSCs governs HF regeneration after injury.

Endogenous ligand promotes hair regeneration via TLR2 on HFSCs

One of the most important endogenous ligands for TLR2 is CEP, which is a naturally occurring product of PUFA oxidation shown to be accumulated during inflammation and wound healing (West et al., 2010, Xiong et al., 2022a). Healthy tissues are typically devoid of this product, which is mainly associated with inflammation and pathologies (Yakubenko et al., 2018). However, in contrast to other tissues, healthy HFs exhibited high levels of CEP accumulation (Fig. 6A). During anagen, CEP is present within the proximal part of the follicle, while in telogen the entire follicle is encased by this PUFA metabolite (Fig. 6B-6E). Generation of CEP from PUFA is directly aided by myeloperoxidase (MPO) (Yakubenko et al., 2018, Xiong et al., 2022a). MPO is present in abundance in sebaceous glands, possibly as a part of immune defense (Supplementary fig. 5A). Even more surprising, in contrast to other organs and tissues, CEP in HFs is substantially depleted with age (Fig. 6F, 6G), and this decline coincides with the reduction in the regenerative potential of HFs. This is likely due to decreased level of MPO during aging (Supplementary fig. 5B, 5C).

Oxidation-dependent TLR2 ligand CEP is present in hair follicles and promotes hair regeneration via HFSC TLR2.

A. Representative images of H&E and CEP immunostaining of consecutive skin sections from WT anagen mouse. Scale bars are 1 mm.

B. Representative confocal images of P5 WT whole-mount skin immunostained for CEP and Ker17. The merged image shows the co-localization of CEP to anagen hair follicles (Ker17+). Scale bar is 200 μm.

C. Longitudinal and cross-sections of anagen and telogen hair follicles from WT mice immunostained for CEP and Ker17. The lower left panel shows a magnified view of the boxed area. Scale bars are 100 μm for anagen, 50 μm for telogen.

D. Quantification of CEP fluorescent intensity at a different distance from the root of anagen hair follicles in longitudinal and cross-sections immunostaining images in images from C. A gradual decrease in CEP levels is observed from the proximal to the distal part of anagen hair follicles. N= 50 follicles from 3 per group.

E. Line chart showing a sharp decrease of CEP fluorescent intensity with the distance from HF in telogen (from the lower right panel in C). N= 10 follicles from 3 per group.

F. Representative confocal images of telogen hair follicles from young and old mice immunostained for CEP and Ker17. Scale bars are 20 μm.

G. Quantification of CEP fluorescent intensity from images in F. N = 6 mice per group.

H. Representative photographs of dorsal skin (two left panels) and inner skin flaps (two right panels) from WT and TLR2KO mice after irradiation and bone-marrow transplantation of WT bone marrow demonstrate an increased number of pigmented hair bulbs and skin pigmentation around wounds in CEP-treated wounds compared to control in WT mice with no differences in TLR2KO transplanted with WT bone marrow. Scale bars are 1 mm for the dorsal skin and 500 μm for the inner skin flap.

I. Quantitative results from H show an increased density of hair follicles upon CEP application around wounds of WT> WT transplanted mice with no changes in WT> TLR2KO mice. N = 5 for each group.

J. Representative photographs of dorsal skin (upper panels) and inner skin flaps (lower panels) from TLR2lox/lox and TLR2HFSC-KO mice treated with CEP show a lack of pigmentation around TLR2HFSC-KO wounds compared with TLR2lox/lox wounds treated with CEP. The inner skin flap of TLR2HFSC-KO mice demonstrates an absence of pigmented hair bulbs after the CEP treatment Scale bars are 3 mm.

K. Representative confocal images of skin adjacent to wound immunostained for Ker17. Scale bars are 100 μm.

L. Quantification of hair follicle numbers in images from K reveals a significant decrease in regenerated hair follicles in TLR2HFSC-KO skin compared with TLR2lox/lox skin.

N=7 for TLR2lox/lox N = 4 for TLR2HFSC-KO.

M. Representative confocal images of skin adjacent to wound immunostained for Ki67. Scale bars are 50 μm.

N. Bar graph showing Ki67 fluorescent intensity in the skin adjacent to wound from images in M. N=7 for TLR2lox/lox N = 4 for TLR2HFSC-KO.

O. Representative microphotographs of primary keratinocytes isolated from WT or TLR2KO mouse skin co-cultured with CEP or control (PBS or BSA). Representative images from at least three independent assays are shown. Scale bar 50 µm.

P. Cell proliferation of primary keratinocytes in O indicates increased proliferation by CEP in WT but not in TLR2KO keratinocytes. N=3 independent experiments.

Q. QPCR analyses of Nfkb2, Il1b, and Il6 mRNA levels in FACS-purified mouse HFSCs treated with BSA control or CEP. N = 3 per group.

R. QPCR analyses of Bmp7 mRNA levels in FACS-purified mouse HFSCs treated with BSA control or CEP. N = 3 per group.

S. Summary of the main findings of this study.

Unpaired t-test (G, P) or Mann-Whitney test (I, L, N, Q, R) was used to determine the statistical significance. All data are mean ± s.e.m.

The connection between CEP levels and hair thinning and loss in aging prompted us to test whether exogenous CEP can activate TLR2 in HFSCs and stimulate their proliferation. Our in-vitro experiments revealed that CEP increases the proliferation of human HFSC in a TLR2-dependent manner since the blockade of TLR2 abrogates the CEP effect (Supplementary fig. 5D, 5E). In another model, CEP promotes cell proliferation of human hair follicle dermal papilla cells (HFDPC) by ∼2-fold compared to control (Supplementary fig. 5F).

Next, we show that CEP promotes hair regeneration in injury in a TLR2-dependent manner. CEP administration promoted HF regeneration in WT wounds. However, it was ineffective in global TLR2KO mice (Supplementary fig. 5H). CEP promoted a 55% increase in the number of HF and cell proliferation in WT wounds, and at the same time, there was no effect in TLR2KO wounds (Supplementary fig. 5I-5J).

To ensure independence from immune cells, WT and TLR2KO mice were irradiated and transplanted with WT bone marrow prior to wounding. Applying CEP on wounds in WT/WT chimeras promoted cell proliferation, thereby dramatically increasing the density of hair follicles (Fig.6H, 6I, Supplementary fig. 5G). At the same time, CEP was not effective in TLR2KO/WT mice (Fig.6H, 6I, Supplementary fig. 5G), demonstrating the TLR2-dependent mechanism.

These CEP effects were mediated by TLR2 on HFSCs. In control mice, CEP effectively initiated regeneration of HFs around the wound (Fig.6J), resulting in ∼3-fold higher density of HF (Fig.6K, 6L) and dramatic acceleration of cell proliferation by >10 fold (Fig.6M, 6N) as compared to TLR2HFSC-KO mice where CEP was mainly ineffective. A similar stimulatory effect of CEP was observed in a primary keratinocyte culture (Oshimori and Fuchs, 2012). CEP dramatically promoted WT but not TLR2KO keratinocyte proliferation (Fig. 6O, 6P). CEP was an effective stimulator of TLR2 signaling as judged by augmented Nfkb2, Il1β, and Il6 expression in HFSCs upon treatment with CEP (Fig. 6Q). Consistent with the key role of BMP signaling in TLR2-dependent HF regeneration, CEP treatment suppressed inhibitory Bmp7 expression by ∼2.5 fold (Fig. 6R), demonstrating that endogenous and natural TLR2 ligand can counteract an inhibitory effect of BMP7 to stimulate HFSCs activation (Fig. 6S).

Discussion

The main findings of this study are as follows: 1) Expression of TLR2 in HFSCs is decreased with aging and in a mouse model of obesity. 2) In young and healthy animals, TLR2 expression in HFs is cycle-dependent, with the highest expression in HFSCs during the initiation of the anagen phase. 3) The absence of TLR2 in HFSCs prolongs the resting phase of the hair cycle and significantly delays hair regeneration after injury. 4) TLR2 regulates the hair cycle primarily by inhibiting BMP signaling in HFSCs. 5) HFs continuously produce a metabolite of PUFAs, which acts as an endogenous TLR2 ligand and promotes hair growth through TLR2 activation in HFSCs. Besides reduced TLR2, aging is linked to low levels of its ligand in HFs. The stimulatory role of TLR2 signaling in HFs was demonstrated through both animal models and established human cell lines.

The lack of TLR2 appears to shift the balance between activating and inhibitory cues, leading to a resting phase that is approximately three times longer. This is a substantial impact considering that there are only 4-5 hair cycles in a mouse lifetime (Choi et al., 2021). The decrease in both TLR2 and its ligand observed in aging and associated conditions will inevitably impede the cyclic regeneration of hair follicles.

The immune system was shown to play a role in the activation of HFSCs, even in the absence of inflammation (Ali et al., 2017, Castellana et al., 2014, Pinho and Frenette, 2019). TLR2 expression increases at the onset of anagen when the immune response is reduced (Paus et al., 2003) and the hair follicle is most susceptible to pathogens. The upregulation of TLR2 in anagen may initially have a protective function. However, high TLR2 expression in undifferentiated vs. differentiated cells underscores its role in stem cell biology. TLRs, including TLR2, have been shown to play a critical role in stem cell functions in various organs (Lathia et al., 2008, Tomchuck et al., 2008, Trowbridge and Starczynowski, 2021). TLRs’ ligation and signaling can alter stem cell differentiation patterns (Collins et al., 2021, Nagai et al., 2006, Trowbridge and Starczynowski, 2021). Proinflammatory signaling can also activate HFSC proliferation, e.g., during injury (Wang et al., 2017, Chen et al., 2015). During the immune-privileged anagen phase of the hair cycle, TLR2 signaling may act as a key intrinsic factor in triggering HFSC activation.

The role of innate immunity in stem cell activation has mainly been linked to TLR3, shown to induce pluripotency in somatic cells through nuclear reprogramming (Lee et al., 2012, Sayed et al., 2015) and drive hair follicle (HF) neogenesis after tissue damage (Nelson et al., 2015). In contrast, we show that TLR2 drives a rapid inflammatory response and regulates the normal hair cycle and HF regeneration/neogenesis in injury.

TLR2 promotes the hair cycle by inhibiting the BMP pathway, a key regulator of hair follicle stem cell (HFSC) quiescence (Plikus et al., 2008, Hsu et al., 2011, Kandyba et al., 2013). Our study demonstrates that reducing excessive BMP signaling reactivates TLR2-deficient HFSCs, revealing a novel link between TLR2, BMP signaling, and the hair cycle. The only known instance of immune system-mediated BMP pathway inhibition occurs during the apoptosis of bulge-associated macrophages (Castellana et al., 2014, Wang et al., 2017).

The role of TLR2 in hair follicle regeneration in both normal hair growth and wound healing emphasizes the importance of understanding the nature of TLR2 ligands mediating these responses. At the site of injury, TLR2 can be activated by pathogens or by endogenously produced ligands, such as the oxidative product of PUFA, CEP, generated in abundance during wound healing (West et al., 2010, Yakubenko et al., 2018). CEP and TLR2 are both essential for hair regeneration, and their deficiency observed in pathologies such as aging and obesity might substantially impair hair growth. Exogenous application of CEP accelerates both wound closure (West et al., 2010) and HF regeneration through TLR2. In addition, both TLR2 and the application of CEP diminish inhibitory BMP signaling, suggesting that CEP and other TLR2 ligands could have therapeutic value for the treatment of hair loss related to burns, traumas, and other pathologies. It is intriguing that while CEP is almost exclusively generated at sites of injury and inflammation (Yakubenko et al., 2018, Xiong et al., 2022a), HFs continuously produce it, most likely by the means of MPO, an anti-bacterial enzyme, capable of generating CEP (Klebanoff et al., 2013).

Contrary to the trend observed in other tissues, where the accumulation of oxidation-generated CEP increases with aging (West et al., 2010), hair follicles (HFs) seem to exhibit depletion of CEP with aging. This decline in CEP levels might contribute to the reduced activity of hair follicle stem cells (Fuchs and Blau, 2020).

The role of CEP in TLR2-dependent HF growth and regeneration highlights the connection between oxidative stress and regenerative processes. Sustained reactive oxygen species (ROS) play a crucial role in proper regeneration, as seen in the tail amputation of Xenopus tadpole (Love et al., 2013). ROS enhance the differentiation of hematopoietic progenitors in Drosophila (Owusu-Ansah and Banerjee, 2009) and sustain self-renewal in neural stem cells (Le Belle et al., 2011). Additionally, ROS production in the skin has been linked to the activation of hair follicle stem cells (HFSCs) (Carrasco et al., 2015). We show the underlying mechanism for these observations, where oxidation-generated CEP triggers TLR2 activation, decreases inhibitory BMP signaling, and stimulates hair follicle growth and regeneration. TLR2 appears to serve as a common link between oxidative stress and tissue regeneration.

To summarize, our study highlights a novel role of TLR2 in promoting tissue regeneration during normal hair growth and wound healing. The identification of an endogenous TLR2 ligand produced by hair follicles presents a potential target for augmenting hair regeneration in the context of injury and aging, opening up new avenues for regenerative medicine.

Supporting information